### 2014-08-09 ms.tex
starting last pass of revisions before submitting to journal

\documentclass{emulateapj}

\usepackage[dvipdf]{epsfig} 
\usepackage[dvips]{rotating}
\usepackage{subfigure}
\usepackage{mathrsfs}
\usepackage{amsmath}

\newcommand{\IRAS}{{\it IRAS}}
\newcommand{\HERSCHEL}{{\it Herschel}}
\newcommand{\SPITZER}{{\it Spitzer}}
\newcommand{\PLANCK}{{\it Planck}}
\newcommand{\AKARI}{{\it Akari}}
\newcommand{\COBE}{{\it COBE}}
\newcommand{\WISE}{{\it WISE}}

\bibpunct{(}{)}{;}{a}{}{,}

\shorttitle{\PLANCK~Two-component Dust Model}

\shortauthors{Meisner \& Finkbeiner}

\begin{document}

\title{Modeling thermal dust emission with two components: application to the 
{\it PLANCK} HFI maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

\begin{abstract}

We apply the \cite{FDS99} two-component thermal dust emission model to the
\PLANCK~HFI maps. This parametrization of the far-infrared dust spectrum as the
sum of two modified blackbodies serves as an important alternative to the 
commonly adopted single modified blackbody (MBB) dust emission model. Analyzing
the joint \PLANCK/DIRBE dust spectrum, we show that two-component models 
provide a better fit to the 100-3000 GHz emission than do single-MBB models, 
though by a lesser margin than found by \cite{FDS99} based on FIRAS and DIRBE. 
We also derive full-sky $6.1'$ resolution maps of dust optical depth and 
temperature by fitting the two-component model to \PLANCK~217, 353, 545, and 
857 GHz along with  DIRBE/\IRAS~100$\mu$m data. Because our two-component model
matches the dust spectrum near its peak, accounts for the flattening of the 
dust spectrum at millimeter wavelengths, and specifies dust temperature at 
6.1$'$ FWHM, our model provides reliable, high-resolution thermal dust emission
foreground predictions from 100 GHz to 3000 GHz. We find that, in diffuse sky 
regions, our two-component 100-217 GHz thermal dust emission predictions are on
average accurate to within 2.2\%, while the \cite{planckdust} single-MBB model 
systematically underpredicts emission by 18.8\% at 100 GHz, 12.6\% at 143 GHz 
and 7.9\% at 217 GHz. We calibrate our two-component optical depth to reddening
at visible wavelengths, and compare the resulting reddening estimates with 
those based on stellar spectra. We find imperfect zodiacal light subtraction in
both DIRBE/\IRAS~and \PLANCK~to be the dominant large angular scale systematic 
problem in our two-component temperature map and reddening estimates.

% might consider changing the word `specifies' in penultimate sentence

% specify early on that this is 'Galactic dust', 'dust' is kind of a vague
% term ... on the other hand the Planck dust paper is pretty lax about this

% mention more explicitly the enhanced temperature resolution relative to SFD??

\end{abstract}

\section{Introduction}
The presence of Galactic interstellar dust affects
astronomical observations over a wide range of wavelengths. In the mid-infrared
and far-infrared, Galactic dust emission contributes significantly
to the total observed sky intensity. At optical and ultraviolet (UV) 
wavelengths, dust grains absorb and scatter starlight. Observations of 
interstellar dust emission/absorption can improve our understanding of the 
physical conditions and composition of the interstellar medium (ISM), an 
environment which plays a crucial role in Galactic evolution and star 
formation. Equally, or perhaps even more important to the practice of 
astronomy, however, is accurately accounting for dust as a foreground which 
reddens optical/UV observations of stars/galaxies and superimposes Galactic 
emission on low-frequency observations of the cosmic microwave background 
(CMB).

% mention re-radiation somewhere ?

Over the past decades, satellite observations have dramatically enhanced our
knowledge about infrared emission from the ISM. The \textit{Infrared Astronomy 
Satellite} (\IRAS), with its $\sim$4$'$ resolution, revolutionized the study of
 Galactic dust emission, first revealing the high-latitude ``infrared cirrus'' 
using 60$\mu$m and 100$\mu$m observations \citep{low84, wheelock94}. Later, the
Diffuse Infrared Background Experiment (DIRBE) aboard the \COBE~satellite 
provided complementary full-sky measurements at ten infrared wavelengths from 
1.25$\mu$m to 240$\mu$m, boasting a reliable zero point despite inferior 
$\sim$0.7$^{\circ}$ angular resolution \citep{boggess92}. \COBE/FIRAS 
\citep{firas} also provided full-sky infrared dust spectra at $7^{\circ}$ 
resolution in 213 narrow frequency bins between 30 GHz and 2850 GHz.

% DIRBE beam solid angles:
% http://lambda.gsfc.nasa.gov/product/cobe/dirbe_products.cfm

\citet[hereafter FDS99]{FDS99} used these FIRAS data to derive a globally 
best-fit model of dust emission applicable over a very broad range of 
frequencies. FDS99 showed that no model consisting of a single modified 
blackbody (MBB) could accurately match the FIRAS/DIRBE spectrum at both the 
Wien and Rayleigh-Jeans extremes. To fit the thermal dust spectrum between 100 
and 3000 GHz, FDS99 therefore proposed an emission model consisting of two 
MBBs, each with a different temperature and emissivity power law index. 
Physically, these two components might represent distinct dust grain species 
within the ISM, or they might simply provide a convenient fitting function. By 
combining this best-fit two-component model with a custom reprocessing of 
DIRBE and \IRAS~100$\mu$m data, FDS99 provided widely used foreground 
predictions with $6.1'$ FHWM, limited largely by their $1.3^{\circ}$ resolution
DIRBE-based temperature correction.

% additional sentence suggested by doug : 
% The fact that the range of optical
% properties of small silicate particles measured in the lab (Agladze et
% al. 1996) spans the range needed by the FDS99 model makes plausible
% that the distinct species are real.

The \PLANCK~2013 data release \citep{planck2013} represents an important 
opportunity to revisit foreground predictions in light of \PLANCK's superb, 
relatively artifact-free broadband data covering the entire sky and a wide 
range of frequencies. \cite{planckdust} has recently conducted a study modeling
\PLANCK~353 GHz, 545 GHz, 857 GHz and DIRBE/\IRAS~100$\mu$m emission with a 
single-MBB spectrum. Here we investigate the FDS99 two-component dust emission 
model as an alernative parametrization for the dust spectral energy 
distribution (SED) composed  of \PLANCK~High Frequency Instrument (HFI), DIRBE 
and \IRAS~data. In doing so, we obtain maps of dust temperature and optical 
depth, both at $6.1'$ resolution. Because we employ a model that has been 
validated with FIRAS data down to submillimeter (submm) wavelengths, we expect 
our best-fit parameters to be useful in constructing high-resolution 
predictions of dust emission over a very broad range of wavelengths. This 
includes low frequencies ($<$350 GHz), which \cite{planckdust} caution their 
model may not adequately fit, and also measurements near the peak of the dust 
SED, for example \AKARI~140-160$\mu$m \citep{akari}. We also anticipate our 
derived optical depth map will serve as a valuable cross-check for extinction 
estimates based directly upon optical observations of stars 
\citep[e.g.][]{schlafly14} and as a baseline for next-generation dust 
extinction maps incorporating high-resolution, full-sky infrared data sets such
as \WISE~\citep{wright10, meisner14} and \AKARI.

In $\S$\ref{sec:data} we introduce the data used throughout this study. In 
$\S$\ref{sec:prepro} we describe our preprocessing of the \PLANCK~maps to 
isolate thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we 
explain the two-component emission model we apply to the \PLANCK-based dust 
SED. In $\S$\ref{sec:bpcorr}, we discuss the details of predicting 
\PLANCK~observations based on this dust model. In $\S$\ref{sec:global} we 
derive constraints on our model's global parameters in light of the \PLANCK~HFI
maps. In $\S$\ref{sec:fitting} we detail the Markov chain Monte Carlo (MCMC) 
method with which we have estimated the spatially varying parameters of our 
model. In $\S$\ref{sec:ebv} we calibrate our derived optical depth to reddening
at optical wavelengths. In $\S$\ref{sec:lofreq} we evaluate the accuracy of our
two-component thermal dust emission predictions from 100-217 GHz. In 
$\S$\ref{sec:release} we present the full-sky maps of dust temperature and 
optical depth we have obtained, and conclude in $\S$\ref{sec:conclusion}.

\section{Data}
\label{sec:data}
All \PLANCK~data products utilized throughout this work are drawn from the 
\PLANCK~2013 release \citep{planck2013}. Specifically, we have made use 
of all six of the zodiacal light corrected HFI intensity maps
\citep[\texttt{R1.10\_nominal\_ZodiCorrected},][]{planckzodi}. Our 
full-resolution (6.1$'$ FWHM) SED fits neglect the two lowest HFI frequencies, 
100 and 143 GHz, as these have FWHM of 9.66$'$ and 7.27$'$ respectively.
% in last sentence, reference Table 1 ?

To incorporate measurements on the Wien side of the dust emission spectrum, 
we include 100$\mu$m data in our SED fits. In particular, we use the 
\citet[henceforth SFD]{SFD} reprocessing of DIRBE/\IRAS~100$\mu$m, which we 
will refer to as \verb|i100|, and at times by frequency as 3000 GHz. The 
\verb|i100| map has angular resolution of $6.1'$, and was constructed so as to 
contain only thermal emission from Galactic dust, with compact sources and 
zodiacal light removed, and its zero level tied to H\,\textsc{i}. We use the 
\verb|i100| map as is, without any custom modifications.

In some of our FIR dust SED analyses which do not require high angular 
resolution, specifically those of $\S$\ref{sec:global}, $\S$\ref{sec:lores}, 
and $\S$\ref{sec:hier}, we also make use of the SFD reprocessings of DIRBE 
140$\mu$m (2141 GHz) and 240$\mu$m (1250 GHz).

\section{Preprocessing}
\label{sec:prepro}

The following subsections detail the processing steps we have applied to 
isolate Galactic dust emission in the \PLANCK~maps in preparation for SED
fitting. % reference fitting section at end of this sentence ?

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}
We first addressed the CMB anisotropies before performing any of the 
interpolation/smoothing described in $\S$\ref{sec:ptsrc}/$\S$\ref{sec:smth}. 
The CMB anisotropies are effectively imperceptible upon visual inspection 
of \PLANCK~857 GHz, but can be perceived at a low level in \PLANCK~545 GHz, and
are prominent at 100-353 GHz relative to the Galactic emission
we wish to characterize, especially at high latitudes. To remove the CMB 
anisotropies, we have subtracted the Spectral Matching Independent Component 
Analysis \citep[SMICA,][]{smica} model from each of the \PLANCK~maps, 
applying appropriate unit conversions for the 545 and 857 GHz maps with native 
units of MJy/sr. Low-order corrections, particularly our removal of  Solar 
dipole residuals, are discussed in $\S$\ref{sec:zp}.

% NEED TO MENTION SMOOTHING OF THE SMICA MAP TO 6.1 ARMCIN !!!!!!

\subsection{Compact Sources}
\label{sec:ptsrc}
% this compact source section still needs a lot of work
After subtracting the SMICA CMB model, we interpolate over compact sources, 
including both point sources and resolved galaxies. Removing compact sources at
this stage is important as it prevents contamination of compact-source-free 
pixels in our downstream analyses which require smoothing of the \PLANCK~maps. 
SFD carefully removed point sources and galaxies from the \verb|i100| map 
everywhere outside of $|b|$$<$$5^{\circ}$. We do not perform any further 
modifications of the \verb|i100| map to account for compact sources. To mask 
compact sources in the \PLANCK~217-857 GHz maps, we use the SFD compact source 
mask. At 100, 143 GHz we use the compact source masks provided by the 
\PLANCK~collaboration in the file \verb|HFI_Mask_PointSrc_2048_R1.10.fits|. 
Given our pixelization (see $\S$\ref{sec:pix}), 1.56\% of pixels are masked 
at 217-857 GHz (1.05\%, 1.02\% at 100, 143 GHz)

% best to refer to an appendix for details of the point source masking ?

\subsection{Smoothing}
\label{sec:smth}
For our full-resolution model, we wish to simultaneously fit \verb|i100| along 
with the four highest-frequency \PLANCK~bands. To properly combine these maps, 
they must have the same point spread function (PSF). \verb|i100|, with its 
$6.1'$ symmetric Gaussian beam, has the lowest angular resolution of the 
relevant maps. To match PSFs, we have therefore smoothed each of the 
\PLANCK~maps under consideration to \verb|i100| resolution by considering each 
native \PLANCK~map to have a symmetric Gaussian beam and smoothing by the 
appropriate symmetric Gaussian such that the resulting map has a  $6.1'$ FWHM. 
The FWHM values we assign to the native \PLANCK~maps are taken from 
\cite{planckbeam}, and are listed in Table \ref{table:offs}.

\subsection{Molecular Emission}
\label{sec:mole}
Because the FIRAS spectra consist of many narrow frequency bins, FDS99 were
able to discard the relatively small number of frequency intervals contaminated
by strong molecular line emission. Unfortunately, while the \PLANCK~data 
considered in this study are of high angular resolution, the broad 
\PLANCK~bandpasses do not allow us to adopt the same approach as FDS99 in 
dealing with line emission. Instead, we must subtract estimates of the 
molecular line contamination from each \PLANCK~band in order to best isolate 
the thermal continuum we wish to characterize. The most prominent molecular 
line emission in the \PLANCK~bands of interest arises from the three lowest CO 
rotational transitions: J=1$\rightarrow$0 at 115 GHz, J=2$\rightarrow$1 at 230 
GHz and J=3$\rightarrow$2 at 345 GHz, respectively affecting the \PLANCK~100, 
217 and 353 GHz bands. The J=1$\rightarrow$0 line also imparts a signal upon 
\PLANCK~143 GHz, but at a negligible level, $\sim$1000$\times$ fainter relative
to the dust continuum than J=1$\rightarrow$0 at 100 GHz.

To correct for molecular emission, we employ the \PLANCK~Type 3 CO data 
product, which boasts the highest S/N among the available full-sky CO maps 
based on the \PLANCK~HFI and Low Frequency Instrument (LFI) data 
\citep{planckco}. The native angular resolution of the Type 3 CO map is 5.5$'$.
We therefore begin by smoothing the raw Type 3 CO map to match the PSF of the 
smoothed \PLANCK~intensity maps we wish to correct for molecular emission.

We must apply the appropriate unit conversions to the Type 3 CO map before 
subtracting it from the \PLANCK~intensity maps, which have native units of 
$K_{CMB}$ at the frequencies of interest. The Type 3 CO map is provided in 
units of K$_{RJ}$ km/s of J=1$\rightarrow$0 emission. To convert this quantity 
to $K_{CMB}$, we assume that all of the CO emission arises from the $^{12}$CO 
isotope, and derive the \PLANCK-observed CO intensity in units of $K_{CMB}$ as
follows:

\begin{equation}
I_{CO, \nu_i, N, N-1} = I_{3}F_{12CO, \nu_i, N, N-1} R_{N, N-1}
\end{equation}

Where $I_{CO, \nu_i, N, N-1}$ is the intensity in $K_{CMB}$ in \PLANCK~band 
$\nu_i$ due to the CO transition from J=$N$ to J=$(N$$-$1). $I_3$ represents 
the appropriately smoothed Type 3 CO amplitude in  K$_{RJ}$ km/s of 
J=1$\rightarrow$0 emission. The $F_{12CO, \nu_i, N, N-1}$ are conversion 
factors between K$_{RJ}$ km/s and $K_{CMB}$ for particular band/transition 
pairs. The relevant values, calculated with the \textit{Unit Conversion and 
Colour Correction} software utilities (\verb|v1.2|), are:

\noindent
$F_{12CO, 100, 1, 0}$=1.478$\times$10$^{-5}$$K_{CMB}/(K_{RJ} \ km/s)$,

\noindent
$F_{12CO, 217, 2, 1}$=4.585$\times$$10^{-5}$$K_{CMB}/(K_{RJ} \ km/s)$, and

\noindent
$F_{12CO, 353, 3, 2}$=1.751$\times$$10^{-4}$$K_{CMB}/(K_{RJ} \ km/s)$.

\noindent
$R_{N, N-1}$ represents the line ratio of the transition from J=$N$ to 
J=$(N$$-$1) relative to the J=1$\rightarrow$0. Thus, $R_{1,0}$=1, and we 
further adopt $R_{2,1}$=0.595 and $R_{3,2}$=0.297 based on \cite{planckco}. 
These line ratios are assumed to be constant over the entire sky.

Formally, then, the CO contamination in band $\nu_i$ is given by:

% might consider quoting rounder values in text, most precise values are:
% 100 GHz, J=1->0 : 1.478e-5, 217 GHz, J=2->1 : 4.585e-5, 
% 353 GHz, J=3->2 : 1.751e-4

\begin{equation} \label{equ:molcorr}
I_{CO, \nu_i} = \sum\limits_{N} I_{CO, \nu_i, N, N-1}
\end{equation}

It happens that, for each of the \PLANCK~bands in which CO emission is
non-negligible (100, 217 and 353 GHz), only a single $N$ contributes ($N$=1, 
$N$=2 and $N$=3, respectively).

Unfortunately, the Type 3 CO map at $6.1'$ FWHM is rather noisy, and the vast
majority of the sky has completely negligible CO emission. Thus, in order to 
avoid adding unnecessary noise outside of molecular cloud complexes and at high
latitudes, we have zeroed out low-signal regions of the Type 3 CO map. We 
identify  low-signal regions as those with $\mathcal{I}_3$$<$1 K$_{RJ}$ km/s, 
where $\mathcal{I}_3$ is the Type 3 CO map smoothed to 0.25$^{\circ}$ FWHM. As 
a result of this cut, 90\% of the sky remains unaffected by our CO correction, 
particularly the vast majority of the high Galactic latitude sky.

% maybe a final sentence or two warning that, to whatever extent
% CO line ratios vary over the sky and/or other molecules make a substantial
% contribution to the observed Planck intensities, our molecular emission
% correction will be insufficient

% for CO line list see ~/wise/pro/planck_co_lines.pro

\subsection{Zero Level}
\label{sec:zp}

% for quantitative assessment of the zero-point difference due to
% flattening of 857 GHz vs. HI, see 
% ~/Desktop/dust/planck_hi/piecewise.pdf (on my Mac)

% -> at low HI, slope decreases by a factor of 1.88
%    the resulting difference in 857 GHz zero points is 0.373 MJy/sr

Although we wish to isolate and model thermal emission from Galactic dust, the
\PLANCK~maps contain additional components on large angular scales. At each 
frequency, there can exist an overall, constant offset that must be subtracted 
to set the zero level of Galactic dust by removing the mean cosmic IR 
background (CIB), as well as any instrumental offset. Additionally, faint 
residuals of the Solar dipole remain at low frequencies. We will address these 
issues by separately solving two sub-problems: first, fixing the absolute zero 
level of \PLANCK~857 GHz relative to external data, and second fitting 
100-545 GHz offsets and low order corrections by correlating these 
\PLANCK~bands against \PLANCK~857 GHz.

\subsubsection{Absolute Zero Level}
\label{sec:zp_abs}
In \cite{planckdust}, the absolute zero level of thermal dust emission was set 
by requiring that \PLANCK~infrared emission tends to zero when H\,\textsc{i} is
zero, assuming a linear correlation between these two measurements at low 
column density. However, this approach is less than completely satsifying in 
that there appear to be different slopes of \PLANCK~857 GHz versus 
H\,\textsc{i} for different ranges of H\,\textsc{i} intensity. In particular, 
\PLANCK~857 GHz appears to ``flatten out'' at very low H\,\textsc{i}, as shown
in Figure 5 of \cite{planckdust}. More quantitavely, we have found using the 
LAB H\,\textsc{i} data \citep{lab} for $-72$$<$$v_{LSR}$$<$$+25$ km/s that the 
best-fit slope for H\,\textsc{i}$<$70 K km/s is a factor of $\sim$1.9 lower 
than the best fit slope for 110 K km/s $<$H\,\textsc{i}$<$200 K km/s, and as a 
result the implied zero level offsets for \PLANCK~857 GHz differ by $\sim$0.37 
MJy/sr.

% maybe state approximately what 0.37 MJy/sr at 857 GHz equates to in EBV
% --> did a quick calculation and it's only ~0.0118 mags ... thought it
%     would be more ??

\begin{figure}
\begin{center}
\epsfig{file=zp_857.eps, width=3.4in}
\caption{\label{fig:fdsref} Scatter plot of FDS99-predicted 857 GHz thermal
dust emission versus \PLANCK~857 GHz observations, illustrating our absolute 
zero level determination described in $\S$\ref{sec:zp_abs}.}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\epsfig{file=dipole_all_100.eps, width=3.4in}
\caption{\label{fig:dip}  Scatter plots of \PLANCK~100, 143, 217, 353, and 545 
GHz versus \PLANCK~857 GHz. Left: before applying our best-fit zero level 
offsets and additional low-order corrections. Right, top four panels: 
\PLANCK~143-545 GHz after correcting for each band's best-fit offset and 
residual Solar dipole. Bottom right: \PLANCK~100 GHz after applying the 
spherical harmonic corrections of Equation \ref{equ:harm}. The dashed red line 
shows the best-fit linear relationship in all cases.}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
\epsfig{file=harm_100.eps, width=6.5in}
\caption{\label{fig:harm} Summary of low-order corrections at 100 GHz. Left: 
prior to our low-order corrections, a $\sim$17$\mu$K zero level offset is 
present and strong low-order problems reduce the linearity of the 100 GHz 
trend versus 857 GHz. Center: scatter plot of \PLANCK~100 GHz versus 857 GHz 
after applying the best-fit offset and residual Solar dipole corrections 
derived with Equation \ref{equ:dip} to \PLANCK~100 GHz. The correlation is 
strengthened, but remains far less tight than for 143-545 GHz (see right 
column of Figure \ref{fig:dip}, top four rows). Right: after applying the 
spherical harmonic corrections of Equation \ref{equ:harm} to \PLANCK~100 GHz, 
the correlation versus 857 GHz is far more tightly linear than following the 
dipole correction.}
\end{center}
\end{figure*}

Because of these ambiguities in the relationship between 857 GHz and 
H\,\textsc{i} emission, we decided to instead constrain the 857 GHz zero level 
by comparison to the FDS99 857 GHz prediction. This renders our \PLANCK~857 GHz
absolute zero level tied indirectly to H\,\textsc{i} through the FDS99 
100$\mu$m and 240$\mu$m zero levels.

We perform a linear fit to the FDS99-predicted 857 GHz values as 
a function of \PLANCK~857 GHz. For this purpose, we employ a version of the 
\PLANCK~857 GHz map with zodiacal light and point sources removed and smoothed 
to 1$^{\circ}$ FWHM, which we will refer to as $\mathcal{I}_{857}$. We 
consider $\mathcal{I}_{857}$ to be the independent variable, as it has much 
higher S/N than the FDS99 prediction, henceforward referred to as 
$\mathcal{F}_{857}$. Note that $\mathcal{F}_{857}$ is not simply the FDS99
model evaluated at 857 GHz, but also incorporates the color correction factor 
of $\S$\ref{sec:bpcorr}, using the FDS99 temperature map to determine the dust 
spectrum shape. We rebin to $N_{side}$=64 and restrict to pixels with 
$\mathcal{I}_{857}$$<$2.15 MJy/sr. Since \PLANCK~857 GHz smoothed to degree 
resolution has very high S/N, we can safely perform such a cut on 
$\mathcal{I}_{857}$. Figure \ref{fig:fdsref} shows a scatter plot of 
$\mathcal{I}_{857}$ versus $\mathcal{F}_{857}$, with a moving median and linear
fit overplotted. The linear fit was performed with uniform weights and 
iterative outlier rejection. The best-fit linear model is given by 
$\mathcal{F}_{857}$=0.991$\mathcal{I}_{857}$$-$0.018 MJy/sr. It is encouraging 
that the slope is quite close to unity.

% i evaluated the `formal statistical error' using the covar keyword in
% hogg_iter_linfit

The formal statistical error on the best-fit 857 GHz offset is quite small, 
$\sim$0.002 MJy/sr. The systematics likely to dominate the actual uncertainty 
on our FDS-based zero level are imperfections in the \PLANCK/\verb|i100| 
zodiacal light models and the FDS99 temperature map. To quantify these 
systematic uncertainties, we split the sky into four quadrants, with boundaries
at $b$=0$^{\circ}$ and $l$=0$^{\circ}$, $l$=180$^{\circ}$. We again restricted 
to $\mathcal{I}_{857}$$<$2.15 MJy/sr, and repeated the regression in each 
quadrant. The rms of the per-quadrant slopes was found to be 0.0188, while the 
rms of the per-quadrant offsets was 0.0586 MJy/sr.

% what HI intensity does the 2.15 MJy/sr 857 GHz correspond to ?
%  ---> pretty high actually

\begin{deluxetable*}{cccccc} 
\tabletypesize{\scriptsize}
\tablecolumns{6} 
\tablewidth{0pc} 
\tablecaption{\label{table:offs} Input Map Properties \& Pre-processing} 
\tablehead{
\colhead{$\nu$ (GHz)} &
\colhead{Offset ($K_{CMB}$)} & 
\colhead{Dipole ($K_{CMB}$)} & 
\colhead{$n_{\nu}$} ($K_{CMB}$) &
\colhead{$c_{\nu}$} &
\colhead{FWHM ($'$)} 
}
\startdata
100  & 1.69$\times$10$^{-5}$$\pm$3.61$\times$10$^{-7}$ & $-$1.08$\times$10$^{-5}$ & 7.77$\times$10$^{-5}$ & 0.0054 & 9.66 \\
143  & 3.58$\times$10$^{-5}$$\pm$7.58$\times$10$^{-7}$ & $-$1.08$\times$10$^{-5}$ & 3.25$\times$10$^{-5}$ & 0.0054 & 7.27 \\
217  & 7.79$\times$10$^{-5}$$\pm$2.60$\times$10$^{-6}$ & $-$1.40$\times$10$^{-5}$ & 4.51$\times$10$^{-5}$ & 0.0054 & 5.01 \\
353  & 2.76$\times$10$^{-4}$$\pm$1.95$\times$10$^{-5}$ & $-$3.08$\times$10$^{-5}$ & 1.51$\times$10$^{-4}$ & 0.012  & 4.86 \\
     & Offset (MJy/sr)                                 & Dipole (MJy/sr)          & $n_{\nu}$ (MJy/sr)    &        &      \\ \cline{2-4} \\ [-2ex]
545  & 7.27$\times$10$^{-2}$$\pm$1.99$\times$10$^{-2}$ & 1.63$\times$10$^{-2}$    & 0.046                 & 0.10   & 4.84 \\
857  & 1.82$\times$10$^{-2}$$\pm$6.02$\times$10$^{-2}$ & -                        & 0.046                 & 0.10   & 4.63 \\
3000 & 0.0$\pm$4.3$\times$10$^{-2}$                    & -                        & 0.06                  & 0.10   & 6.1
\enddata
\end{deluxetable*}

\subsubsection{Relative Zero Level}
\label{sec:relzero}

In the course of this study we use not only \PLANCK~857 GHz, but also all of
the remaining \PLANCK~HFI bands, as well as \verb|i100|. To derive the 
zero level offsets that must be applied to each of the five lowest-frequency 
\PLANCK~bands, we perform a regression versus the \PLANCK~857 GHz map corrected
for the best-fit absolute zero level offset from $\S$\ref{sec:zp_abs}. We 
assume no offset need be applied to \verb|i100|, which already has its zero 
level tied to H\,\textsc{i}.

The need for additional low-order corrections beyond simple scalar offsets 
became evident upon inspecting the HFI maps at 100-545 GHz. In particular, we 
noticed the presence of a low-level dipole pattern, with an orientation 
consistent with that of the Solar dipole. Our strategy will be to 
simultaneously fit both this residual dipole and the zero-level offset 
amplitude for each band. To most precisely recover these amplitudes, it is 
necessary to have the highest available S/N in the independent variable of our 
regression. For this reason we have used \PLANCK~857 GHz as a reference for the
100-545 GHz bands, as opposed to the FDS99 predictions or H\,\textsc{i} data. 
In doing so, we assume \PLANCK~857 GHz contains no appreciable Solar dipole 
residual.

We perform one regression per HFI band (other than 857 GHz) to simultaneously 
fit for the zero level offset, the slope relative to 857 GHz, and the residual 
dipole amplitude. For each 100-545 GHz HFI band, we restrict to regions of low 
column density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for 
$-72$$<$$v_{LSR}$$<$$+25$ km\,s$^{-1}$) and fit the following model:

\begin{equation} \label{equ:dip}
\mathcal{I}_{\nu_i, p} = m\mathcal{I}_{857, p} + b + d\mathcal{D}_{p}
\end{equation}

With $p$ denoting a single $N_{side}$=64 HEALPix pixel \citep{healpix} in the 
maps $\mathcal{I}_{857}$, $\mathcal{I}_{\nu_i}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and the constant offset of $\S$\ref{sec:zp_abs} removed, smoothed to 
$1^{\circ}$ resolution. $\mathcal{I}_{\nu_i}$ is the corresponding $1^{\circ}$ 
resolution \PLANCK~HFI map with zodiacal emission, CMB anisotropies, and
compact sources removed. In the context of Equation \ref{equ:dip}, $\nu_i \in$ 
\{100, 143, 217, 353, 545\} GHz. Note that $\mathcal{I}_{\nu_i}$ is always in
the native units of the relevant \PLANCK~band. $\mathcal{D}$ is a scaling of 
the Solar dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$, with unit amplitude. Because 
$\sim$18,000 pixels satisfy the low H\,\textsc{i} cut, we have an 
overconstrained linear model with three parameters: $m$, $d$, and $b$. $m$ 
represents the best-fit slope of \PLANCK~band $\nu_i$ versus \PLANCK~857 GHz 
assuming they are linearly related. $d$ is the residual Solar dipole amplitude,
and its best-fit value represents the scaling of the Solar dipole that makes 
the \PLANCK~band $\nu_i$ versus 857 GHz correlation most tightly linear. $b$ 
represents the constant offset that must be subtracted from the band $\nu_i$ 
map to make its zero level consistent with that of the 857 GHz map. For each 
band $\nu_i$, we obtain estimates of $m$, $d$, and $b$ by performing a linear 
least squares fit with uniform weights and iterative outlier rejection. Figure 
\ref{fig:dip} shows scatter plots of the band $\nu_i$ versus 857 GHz 
correlation before and after correcting for the best-fit offset and residual 
dipole, for each $\nu_i \in$ \{143, 217, 353, 545\} GHz. Not only are the 
tightened correlations striking in these scatter plots, but the residual dipole
subtractions appear very successful in the two-dimensional band $\nu_i$ maps 
themselves. Before performing thermal dust fits, we therefore subtract the 
best-fit $b$ and $d\mathcal{D}$ from each band $\nu_i$ map. The best-fit 
offsets and residual dipole amplitudes are listed in Table \ref{table:offs},
along with other important per-band parameters, such as the fractional 
multiplicative calibration uncertainty $c_{\nu}$.

We found that a dipole correction alone could not sufficiently rectify the 
\PLANCK~100 GHz map (see Figure \ref{fig:harm}). Therefore, for 100 GHz, we 
performed a modified version of the Equation \ref{equ:dip} fit, using the 
following model:

\begin{equation} \label{equ:harm}
\mathcal{I}_{100, p} = m\mathcal{I}_{857, p} + b + \sum_{l=1}^{4} \sum_{m=-l}^{l} a_{lm}Y_{l}^{m}(\theta_p, \phi_p)
\end{equation}

Where $Y_{l}^{m}$ are the real spherical harmonics, and the $a_{lm}$ are their 
corresponding real coefficients. The angle $\phi_p$ is taken simply to be 
$l_{gal, p}$ and $\theta_p$=$(90^{\circ}-b_{gal, p})$. Thus, we have replaced 
the Solar dipole term with a sum of 24 spherical harmonic templates, which, 
when multiplied by the best-fit $a_{lm}$ coefficients and subtracted from 
\PLANCK~100 GHz make the relation between 100 GHz and 857 GHz most tightly 
linear. Figure \ref{fig:harm} illustrates the improved correlation of 100 GHz 
vs. 857 GHz when including the spherical harmonic corrections relative to the 
dipole-only correction. The spherical harmonic decomposition of Equation 
\ref{equ:harm} did not improve the correlations at higher frequencies 
enough to warrant replacing the dipole-only correction in those cases.

% mention how offset/dipole values compare to those derived in the Planck
% team paper ?

\section{Dust Emission Model}
\label{sec:modeling}

% maybe a sentence to transition from preprocessing -> modeling
% e.g. With pre-processing completed, we turned to modeling the 
% thermal dust emission.

At sufficiently high frequencies, Galactic thermal dust emission can be 
adequately modeled as a single MBB with power-law emissivity 
\citep[e.g. SFD;][]{planckdust}. However, it has long been recognized, 
particularly in view of the FIRAS spectra, that the dust SED flattens toward 
the millimeter in a manner which is not consistent with a simple extrapolation 
of the single-MBB model to low frequencies. In the diffuse ISM, \cite{reach95} 
found an improved fit to the FIRAS data using an empirically motivated 
superposition of two $\beta$=2 MBBs, one representing a `hot' grain population 
($T$$\approx$16$-$21 K), the other a `cold' grain population 
($T$$\approx$4$-$7 K). FDS99 built a more physically motivated two-MBB model, 
in which different grain emission/absorption properties account for the 
differing temperatures of each population, and these temperatures are coupled 
by assuming thermal equilibrium with the same interstellar radiation field 
(ISRF).

The primary FDS99 analysis considered the intrinsic grain properties of each
species, for example the emissivity power law indices, to be constant over the 
sky, and performed a correlation slope analysis to constrain these parameters 
with FIRAS and DIRBE observations. FDS99 also constructed a DIRBE 
240$\mu$m/100$\mu$m ratio to account for temperature variation at 
$\sim$1.3$^{\circ}$ resolution. In this work we seek to apply the FDS99 
emission model to the \PLANCK~data set, which offers a dramatic enhancement in
angular resolution relative to the FIRAS spectra. The \PLANCK~data thereby 
allow us to derive an improved temperature correction at near-\IRAS~resolution
($\S$\ref{sec:mcmc}), re-evaluate the best-fit global dust properties 
($\S$\ref{sec:global}, $\S$\ref{sec:hier}), and fit additional 
two-component model parameters as a function of position on the sky 
($\S$\ref{sec:lores}).

% kind of feel like i'm re-writing the introduction here ??

% could add a last sentence 'ultimately this allows us to make high-resolution
% foreground and extinction predictions' and/or even find a way to mention
% other predictions, such as polarization/R_v

% in the above paragraph, should I mention the Planck collaboration result
% of beta_mm vs. beta_fir being different, with (beta_fir-beta_mm) > 0

% should i also try to justify why it's worth doing this fit even
% if it's possible to just claim different beta's at different frequencies?

The shape of the two-component model spectrum we will consider is given by:

\begin{equation}
M_{\nu} \propto \Big[f_{1}q_{1}\Big(\frac{\nu}{\nu_{0}}\Big)^{\beta_1}B_{\nu}(T_1) + f_{2}q_{2}\Big(\frac{\nu}{\nu_0}\Big)^{\beta_2}B_{\nu}(T_2)\Big]
\end{equation}

% With $f_{1}$ = 0.0363, $\beta_1$ = 1.67, $\beta_2$ = 2.70, and $q_1/q_2$ = 
% 13.0.

% re: following paragraph --- in the code nu_0 is actually 2997.92458
% because this is the frequency with respect to which DIRBE color correction
% K_100 (and my color correction as well) is referenced

% not sure if `Planck function' is the correct jargon

Where $B_{\nu}$ is the Planck function, $T_1$ is the `cold' dust temperature, 
$T_2$ is the `hot' dust temperature, and $\beta_1$ and $\beta_2$ are the
emissivity power-law indices of the cold and hot dust components respectively. 
$q_1$ represents the ratio of FIR emission cross section to optical absorption 
cross section for species 1, and similarly $q_2$ for species 2. $f_1$ and 
$f_2$ dictate the relative contributions of the two MBB components to the
combined SED, and can be thought of as encoding the mass fraction of each 
species, although technically $f_1$ ($f_2$) is the optical absorption 
cross-section weighted mass fraction for species 1 (2). Following the 
convention of FDS99, we choose $\nu_0$=3000 GHz and take $f_2$=(1$-$$f_1$).

Mathematically, this two-MBB model requires specification of seven 
parameters for every line of sight: $T_1$, $T_2$, $\beta_1$, $\beta_2$, $f_1$, 
$q_1$/$q_2$ and the normalization of $M_{\nu}$. However, under the 
assumption that the temperature of each species is determined by maintaining 
thermal equilibrium with the same ISRF, $T_1$=$T_1$($T_2$, $\beta_1$, 
$\beta_2$, $q_1/q_2$) is fully determined by these other parameters. $T_1$ is 
always related to $T_2$ via a simple power law, although the prefactor and 
exponent depend on the parameters $q_1/q_2$, $\beta_1$ and $\beta_2$ (see FDS99
Equation 14).

% Importantly, the emissivity power law indices $\beta$ of the two species 
% differ.

These considerations still leave us with six potentially free parameters per 
line of sight. Unfortunately, fitting this many parameters per spatial pixel is
not feasible for our full-resolution $6.1$$'$ fits, as these are constrained by
only five broadband intensity measurements. Hence, as in FDS99, we deem certain
 parameters to be ``global'', i.e. spatially constant over the entire sky. In 
our full-resolution five-band fits, we designate $\beta_1$, $\beta_2$, $f_1$ 
and $q_1/q_2$ to be spatially constant. This same approach was employed by 
FDS99, and the globally best-fit values obtained by FDS99 for these parameters 
are listed in the first row of Table \ref{tab:global}. With these global 
parameters, FDS99 found $T_2$$\approx$$16.2$K, $T_1$$\approx$$9.4$K to be 
typical at high-latitude. In $\S$\ref{sec:global}, we discuss the best-fit 
global parameters favored by the \PLANCK~HFI data; these are listed in the 
second row of Table \ref{tab:global}.

%%%%

\begin{deluxetable*}{llrrrrrrrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{11} 
\tablewidth{0pc} 
\tablecaption{\label{tab:global} Global Model Parameters} 
\tablehead{
\colhead{Number} &
\colhead{Model} &
\colhead{$f_1$} &
\colhead{$q_1/q_2$} & 
\colhead{$\beta_1$} & 
\colhead{$\beta_2$} &
\colhead{$T_2$} &
\colhead{$T_1$} &
\colhead{$n$} &
\colhead{D.O.F.} &
\colhead{$\chi^2$} &
\colhead{$\chi^2_{\nu}$}
}
\startdata
 1 & FDS99 best-fit  & 0.0363 & 13.0  & 1.67 & 2.70 & 15.72 &  9.15 & 1.018 & 7 & 23.9 & 3.41 \\
 2 & FDS99 general   & 0.0485 & 8.22  & 1.63 & 2.82 & 15.70 &  9.75 & 0.980 & 3 & 3.99 & 1.33 \\
 3 & single MBB      &  0.0   &  ...  &  ... & 1.59 & 19.63 &   ... & 0.999 & 6 & 33.9 & 5.65 \\ [-2ex]
\enddata
\end{deluxetable*}

% would be great to have uncertainties on each parameter as well (in the
% Planck case at least...) !!!!!!!!!!!!!!!

%%%%

Fixing the aforementioned four global parameters, our full-resolution, 
five-band fits have two remaining free parameters per line of sight: the hot 
dust temperature $T_2$ determines the SED shape and the normalization of 
$M_{\nu}$ determines the SED amplitude. In the lower-resolution fits of 
$\S$\ref{sec:lores} which include all HFI bands, we will allow $f_1$ to be a 
third free parameter, still holding $\beta_1$, $\beta_2$, and $q_1/q_2$ fixed.

%explain that we take the ``hot'' dust temperature to be T2 rather than T1

% maybe point to the SED fit summary figure as an example of what
%

To calculate the optical depth in the context of this model, we assume
optically thin conditions, meaning that $\tau_{\nu}$ = $M_{\nu}/S_{\nu}$, where
$M_{\nu}$ is the appropriately scaled two-component model intensity and the 
source function is given by:

\begin{equation}
\label{eqn:source}
S_{\nu} = \frac{f_1q_1(\nu/\nu_0)^{\beta_1}B_{\nu}(T_1) + f_2q_2(\nu/\nu_0)^{\beta_2}B_{\nu}(T_2)}{f_1q_1(\nu/\nu_0)^{\beta_1}+f_2q_2(\nu/\nu_0)^{\beta_2}}
\end{equation}

\section{Predicting the Observed SED}
\label{sec:bpcorr}

The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
flux density per solid angle $M_{\nu}$ in e.g. MJy/sr for any single frequency 
$\nu$. In practice, however, we wish to constrain our model using measurements 
in the broad \PLANCK/DIRBE bandpasses, each with $\Delta\nu/\nu\sim0.3$. 
Both the \PLANCK~and DIRBE data products quote flux density per solid 
angle in MJy/sr under the `IRAS convention'. More precisely, each value 
reported in the \PLANCK~maps gives the amplitude of a power-law spectrum 
with $\alpha$=$-1$, evaluated at the nominal band center frequency, such that 
this spectrum integrated against the transmission reproduces the 
bolometer-measured power. Because our model spectra do not conform to the 
$\alpha$=$-1$ convention, we have computed color correction factors to account 
for the MBB($T$, $\beta$) spectral shape and the transmission as a function of 
frequency:

% is delta(nu)/nu conventionally (nu_off-nu_on)/nu_c or 1/2 of that ?

\begin{equation} \label{equ:bpcorr}
b_{\nu_i}(T, \beta) = \frac{\int \nu^{\beta}B_{\nu}(T)\mathcal{T}_{\nu_i}(\nu) d\nu \bigg[\int (\nu_{i,c}/\nu)\mathcal{T}_{\nu_i}(\nu) d\nu\bigg]^{-1}}{\nu_{i,c}^{\beta}B_{\nu_{i,c}}(T)}
\end{equation}

% integral limits ??? check planck paper to see what they did
% -> planck spectral response paper just leaves the integrals indefinite

Here $\nu_{i,c}$ is the nominal band center frequency of band $\nu_i$,  with 
$\nu_{i,c} \in$ \{100, 143, 217, 353, 545, 857, 1249.1352, 2141.3747, 
2997.92458\} GHz. $\mathcal{T}_{\nu_i}(\nu)$ represents the relative 
transmission as a function of frequency for band $\nu_i$. For the HFI maps, 
$\mathcal{T}_{\nu_i}(\nu)$ is given by the \PLANCK~transmission curves provided
in the file \verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. For \verb|i100| 
and DIRBE 140$\mu$m, 240$\mu$m, we have adopted the corresponding DIRBE 
transmission curves.

%Again, since DIRBE-quoted fluxes also follow the 
%$\alpha$=$-1$ convention, Equation \ref{equ:bpcorr} remains applicable.

% what is the reference for the dirbe

%The color correction factors
%of Equation \ref{equ:bpcorr} are applied in the sense that 
%$b_{\nu_i}(T ,\beta)\times M_{\nu_{i,c}}$ is a prediction suitable for 
%direct comparison to the \PLANCK~data products.

The two-component model prediction in band $\nu_i$ under the IRAS convention, 
termed $\tilde{I}_{\nu_i}$, is then constructed as a linear combination of 
color-corrected MBB terms:

\begin{equation} \label{equ:iras}
\tilde{I}_{\nu_i} \propto \sum_{k=1}^{2} b_{\nu_i}(T_k, \beta_k) f_k q_k (\nu_{i,c}/\nu_0)^{\beta_k} B_{\nu_{i,c}}(T_k)
\end{equation}

%Note that, as in FDS99, $f_2$=(1$-$$f_1$) by convention. 
The color correction of Equation \ref{equ:bpcorr} therefore allows us to 
predict $\tilde{I}_{\nu_i}$ by computing monochromatic flux densities at the 
central frequency $\nu_{i,c}$ and then multiplying by factors 
$b_{\nu_i}(T, \beta)$. In practice, we interpolated the color corrections off 
of a set of precomputed, one-dimensional lookup tables each listing 
$b_{\nu_i}(T, \beta)$ as a function of $T$. We avoided the need to interpolate 
in both $\beta$ and $T$ by computing only a small set of one dimensional 
correction factors for the particular set of $\beta$ values of interest (e.g. 
$\beta$=1.67, 2.7, 1.63, 2.82 ..., see Table \ref{tab:global}). This color 
correction approach makes the MCMC sampling described in $\S$\ref{sec:mcmc} 
much more computationally efficient by circumventing the need to perform the 
integral in the numerator of Equation \ref{equ:bpcorr} on-the-fly for each 
proposed dust temperature. We have chosen to compute the color corrections on a
per-MBB basis because this approach is very versatile; all possible 
two-component (and single-MBB) models are linear combinations of MBBs, so we 
can apply all of our color correction machinery even when we allow parameters 
other than temperature (e.g. $f_1$) to vary and thereby modify the dust 
spectrum shape.

% need some sort of clarification about the fact that, in the parts of the 
% analysis where beta floats, i actually do interpolate color corrections 
% off of a 2d grid in (T, beta)

% We note that it is not clear which bandpass should properly be ascribed to 
% the \IRAS/DIRBE composite 100$\mu$m maps under consideration.

With these color corrections and the  formalism established in 
$\S$\ref{sec:modeling} in hand, we can mathematically state the model we will 
use e.g. during MCMC sampling to predict the observed SED. The predicted 
observation in band $\nu_i$ is given by:

\begin{equation}
\label{eqn:inten}
\tilde{I}_{\nu_i} = \frac{\sum\limits_{k=1}^{2} b_{\nu_i}(T_k, \beta_k) f_k q_k (\nu_{i,c}/\nu_0)^{\beta_k} B_{\nu_{i,c}}(T_k) u_{\nu_i}^{-1}(T_k, \beta_k)}{\sum\limits_{k=1}^{2} b_{545}(T_k, \beta_k) f_k q_k (545 \textrm{GHz}/\nu_0)^{\beta_k} B_{545}(T_k)}\tilde{I}_{545}
\end{equation}

This equation is quite similar to Equation \ref{equ:iras}, but with two 
important differences. First, the normalization of $\tilde{I}_{\nu_i}$ is now 
specified by $\tilde{I}_{545}$, which represents the IRAS convention 
\PLANCK~545 GHz intensity. The denominator serves to ensure that, for the case 
of $\nu_i$=545 GHz, $\tilde{I}_{545}$ is self-consistent. Second, each term in 
the numerator is multiplied by a unit conversion factor 
$u_{\nu_i}^{-1}(T_k, \beta_k)$. This factor is necessary because some of the 
\PLANCK~maps of interest have units of $K_{CMB}$ (100-353 GHz), while the
remaining maps (545-3000 GHz) have units of MJy/sr. We have adopted the 
strategy of predicting each band in its native units, whether MJy/sr or 
$K_{CMB}$. For this reason, we always evaluate $B_{\nu_{i,c}}$ in Equation 
\ref{eqn:inten} in MJy/sr and let $u_{\nu_i}$=1 (dimensionless) for 
$\nu_i$$\ge$545 GHz. For $\nu_i$$\le$353 GHz, $u_{\nu_i}(T_k, \beta_k)$ 
represents the conversion factor from $K_{CMB}$ to MJy/sr, given by 
\cite{planckresponse} Equation 32.

%is the old sentence below worth reinstating ?

%To reiterate, for 
%$\nu_i$$\le$353 GHz, $u_{\nu_i}(T_k, \beta_k)$ has units of 
%(MJy/sr)/$K_{CMB}$,and therefore $\tilde{I}_{\nu_{i}}$ will have different 
%units depending on the band $\nu_i$, though $\tilde{I}_{\nu_{i}}$ always has 
%the same units as the native band $\nu_i$ map.

% check that RIMO file i've used is actually the most up-to-date version

% extremely abrupt transition here...
\section{Global Model Parameters}
\label{sec:global}

\begin{figure*}
\begin{center}
\epsfig{file=dirbe_slopes.eps, width=6.5in}
\caption{\label{fig:dirbe_slopes} Linear fits of SFD-reprocessed DIRBE 
240$\mu$m (left), 140$\mu$m (center), and 100$\mu$m (right) as a function of 
\PLANCK~857 GHz. The red lines illustrate the DIRBE correlation slopes used in 
our dust emission model optimization of $\S$\ref{sec:global}.}
\end{center}
\end{figure*}

While we ultimately aim to obtain \PLANCK-resolution maps of the spatially 
varying dust temperature and optical depth, we start by applying the 
machinery/formalism thus far developed to reassess the best-fit global 
two-component model parameters in light of the \PLANCK~HFI data.

FDS99 determined the best-fit values of the two-component model global 
parameters $\beta_1$, $\beta_2$, $q_1/q_2$ and $f_1$ via a correlation slope
analysis incorporating DIRBE and FIRAS data. Here we seek to estimate these 
same global parameters via an analogous correlation slope analysis in which we 
swap the \PLANCK~HFI maps for FIRAS at low frequencies, while still relying on 
DIRBE at higher frequencies. We also seek to determine via this correlation 
slope analysis whether or not the combination of \PLANCK+DIRBE data favors 
two-component models over single-MBB models in the same way that the 
FIRAS+DIRBE data did in the FDS99 analysis.

In the two-component model case, based on a spectrum of \PLANCK~and 
DIRBE correlations slopes, we wish to obtain estimates for six free parameters:
$\beta_1$, $\beta_2$, $q_1/q_2$, $f_1$, $T_2$ and the overall spectrum 
normalization $n$. The constraints we employ are the correlation slopes of each
of the \PLANCK~HFI bands, as well as DIRBE 100$\mu$m (3000 GHz), 140$\mu$m 
(2141 GHz) and 240$\mu$m (1250 GHz) relative to \PLANCK~857 GHz, i.e. 
$dI_{\nu_i}/dI_{857}$. We will refer to the slope for band $\nu_i$ relative to 
\PLANCK~857 GHz as $s_{i,857}$. The slopes for \PLANCK~100-545 GHz are taken 
to be those derived from the relative zero level fits of $\S$\ref{sec:relzero},
and are illustrated by the dashed red lines in the right-hand column plots of 
Figure \ref{fig:dip}. The 857 GHz slope is unity by definition.

At 1250, 2141 and 3000 GHz, we use the SFD-reprocessed DIRBE maps. For each
DIRBE band, we determine $s_{i, 857}$ by performing a linear fit to DIRBE as a 
function of \PLANCK~857 GHz, after both have been zodiacal light subtracted and
smoothed to $1^{\circ}$ FWHM, also restricting to the low HI mask of 
$\S$\ref{sec:relzero} (see Figure \ref{fig:dirbe_slopes}).
% re-mention nside=64 or just bag it?
% restricted to same low HI mask as in Sect. 3.5.2 analysis
% sentence pointing to figure showing the regressions

Counting 857 GHz, we thus have nine correlation slope constraints for six 
free parameters. Including DIRBE 140$\mu$m and 240$\mu$m is critical in making 
the problem at hand sufficiently overconstrained, and also in providing 
information near the peak of the dust SED at $\sim$160$\mu$m, which is 
particularly sensitive to the presence of a single versus multiple MBB 
components.

We assume an uncertainty of 2\% on each of the $s_{i, 857}$ and minimize the 
chi-squared given by:

\begin{equation} \label{equ:chi2corr}
\chi^2 = \sum_{i=0}^{8}\frac{\big[s_{i, 857}-n\frac{\tilde{I}_{\nu_i}(\beta_1, \beta_2, f_1, q_1/q_2, T_2)}{\tilde{I}_{857}(\beta_1, \beta_2, f_1, q_1/q_2, T_2)}\big]^2}{\sigma_{s_i}^2}
\end{equation}

Where $\nu_i$$\in$\{100, 143, 217, 353, 545, 857, 1250, 2141, 3000\} GHz. Note 
that this formula encompasses the general two-component case; in the single-MBB
case, we take $f_1$=0 and hence $q_1/q_2$, $\beta_1$ and $T_1$ are immaterial, 
but Equation \ref{equ:chi2corr} still applies. Note also that no `priors' are
included to preferentially drag our results towards agreement with those of 
FDS99.

The results of our chi-squared minimization are listed in Table 
\ref{tab:global}. First (model 1), we fix $\beta_1$, $\beta_2$ ,$q_1/q_2$ and
$f_1$ to the best-fit values from the FDS99 analysis based on DIRBE+FIRAS. We
then allow $n$ and $T_2$ to vary so as to best match our 
DIRBE+\PLANCK~spectrum. This results in a reduced chi-squared of 
$\chi^2_{\nu}$=3.41. Reassuringly, $n$ is quite close to unity. It should be 
noted though that our best-fit $T_2$ is $\sim$0.5 K lower than that found by 
FDS99 for the same values of $\beta_1$, $\beta_2$, $q_1/q_2$ and $f_1$.

Next (model 2), we consider the fully-general two-component model, allowing all
six model parameters to vary. In this case, the reduced chi-squared of the best
fit parameters is $\chi^2_{\nu}=$1.33, signifying that our introduction of four
additional free parameters is justified. The best-fit $\beta_1$ and $\beta_2$ 
are both consistent with the corresponding FDS99 values to within 5\%. 
$q_1/q_2$=8.22 represents a $\sim$40\% lower value than found by FDS99, while
$f_1$=0.0458 represents a $\sim$25\% increase relative to FDS99. Again, our
best-fit high-latitude $T_2$ is $\sim$0.5 K lower than the typical value of 
$\langle T_2 \rangle$=16.2 K from FDS99.

Lastly, we calculate the optimal single-MBB fit to the \PLANCK+DIRBE 
correlation slope spectrum. The best-fit single MBB has $\beta$=1.59, 
$T$=19.63, and $\chi^2_{\nu}$=5.65, indicating a significantly worse fit to the
data than our best-fit two-component model (model 2). Thus, our \PLANCK+DIRBE 
correlation slope analysis has confirmed the main conclusion of FDS99 and 
others e.g. \cite{reach95}, that the FIR/submm dust SED prefers two MBBs to 
just one, but, for the first time, independent of FIRAS. Still, it is apparent 
that the improvement in $\chi^2_{\nu}$ for single-MBB versus double MBB models 
found here is substantially less dramatic ($\Delta\chi^2_{\nu}$=4.32) than that
found in FDS99 ($\Delta\chi^2_{\nu}$=29.2). This is likely attributable to the 
exquisite narrow-band frequency coverage of FIRAS, especially near the dust SED
peak, which makes FIRAS a better suited data set than \PLANCK~for a detailed 
analysis of the globally best-fit dust SED model. In $\S$\ref{sec:hier}, we
confirm the basic conclusions of this section via an approach in which we 
allow the dust temperature to vary spatially. The analysis of $\S$7.5 also
allows us to confirm the conclusions of this section while including a 
fully detailed uncertainty model; our assumption of 2\% per-band uncertainties
on the correlation slopes is largely a statement that we seek a model
accurate to 2\% from 100-3000 GHz, although the fact that our $\chi^2_{\nu}$ 
values are order unity shows that the assumed uncertainties are not grossly 
over or underestimated.

% maybe a table listing the correlation slopes and their uncertainties
% would be useful here

\section{MCMC Fitting Procedure}
\label{sec:fitting}

The following subsections detail our procedure for constraining the 
two-component dust emission model parameters which are permitted to vary
spatially. We use the MCMC procedure described to perform two types of fits: 
(1) full-resolution $6.1'$ fits, in which only  the SED normalization and dust 
temperatures vary spatially, and (2) lower-resolution fits in which $f_1$ is 
also allowed to vary from one line of sight to another.

\begin{figure}
\begin{center}
\epsfig{file=sed.eps, width=3.3in}
\caption{\label{fig:sed} Observed SED (black points and error bars) and 
best-fit thermal dust emission models for a single $N_{side}$=2048 pixel in the
Polaris flare region. The best-fit two-component model (black line) is derived 
based on \PLANCK~217-857 GHz and SFD 100$\mu$m at full $6.1'$ resolution. Note 
that the two lowest-frequency data points were not used in the two-component 
fit shown, while the three lowest-frequency data points were not used in the 
\cite{planckdust} single-MBB fit shown (cyan).}
\end{center}
\end{figure}

% add error bars to this example SED plot

\begin{figure*}
\begin{center}
\epsfig{file=posteriors.eps, width=7.0in}
\caption{\label{fig:post} Gridded posterior PDFs for three $N_{side}$=2048 
pixels, based on \PLANCK~217-857 GHz and SFD 100$\mu$m at full $6.1'$ 
resolution. The colorscale is linear in $log(P)$, with black corresponding to 
the maximum of $log(P)$ and white representing $max[log(P)]-5$. Light green 
crosses and ellipses mark the best-fit parameters and $1\sigma$ uncertainties 
based on our MCMC sampling of the posteriors. Our MCMC parameter and 
uncertainty estimates are in good agreement with those based on gridded 
posteriors. These three pixels are also representative in that we find the 
posterior distributions from Equation \ref{eqn:post} are in general extremely 
well-behaved, showing no multimodality or other pathological qualities. Left: 
Low S/N pixel at high latitude in the Galactic north. Center: High S/N pixel in
the Polaris flare region. Right: Low S/N pixel at high latitude in the Galactic
south.}
\end{center}
\end{figure*}

% overplot ellipses showing 1 sigma errors in posterior grayscales

\subsection{Pixelization}
\label{sec:pix}
For the purpose of fitting, we divide the sky into $\sim$50 million pixels of 
angular size $\sim$1.72$'$, defined by the HEALPix pixelization in Galactic 
coordinates, with $N_{side}$=2048. This pixelization is convenient because it 
is the format in which the \PLANCK~HFI maps were released, and because it 
adequately samples the $6.1'$ FWHM maps under consideration in our 
full-resolution fits. Our procedure will fit the intensity measurements in each
spatial pixel independently.

\subsection{Sampling Parameters}
\label{sec:samp}
As discussed in $\S$\ref{sec:modeling}, our full-resolution fits
consider the ``global'' parameters $f_1$, $q_1/q_2$, $\beta_1$, $\beta_2$ to be
spatially constant. We employ the best-fit \PLANCK+DIRBE global parameters of 
Table \ref{tab:global}, model 2. For each line of sight, only the dust spectrum
normalization and dust temperatures are allowed to vary. In order to predict 
the dust SED for a given pixel, we are thus left with two remaining degrees of 
freedom, and must choose an appropriate set of two parameters to sample and 
thereby constrain via MCMC. To determine the SED normalization in each pixel, 
we draw samples in $\tilde{I}_{545}$, the `IRAS convention' intensity in the 
\PLANCK~545 GHz bandpass, as defined in Equation \ref{eqn:inten}. With the four
aforementioned global parameters fixed, the dust spectrum shape is determined 
entirely by the two dust temperatures, which are coupled. To constrain the dust
temperatures, we sample in $T_2$, the hot dust temperature. For each sample in 
$T_2$, we compute the corresponding value of $T_1$, thereby determining the 
temperature and emissivity power law index for both MBB components and thus 
fully specifying the SED shape. In principle, we could sample in either $T_1$ 
or $T_2$, but have chosen to sample in $T_2$ because emission from this 
component dominates in the relatively high frequency bands which most strongly 
constrain the dust temperatures.

For the lower resolution fits described in $\S$\ref{sec:lores}, we sample
in three parameters: $\tilde{I}_{545}$, $T_2$, and $f_1$.

% sentence or two mentioning reduced resolution fits where we sample in
% three dimensions (f_1, T_2, tilde{I}_{545})

\subsection{Markov Chains}
\label{sec:mcmc}

In our full-resolution fits, we use a MCMC approach to constrain the 
parameters $\tilde{I}_{545}$ and $T_2$. For each pixel, we run a 
Metropolis-Hastings (MH) Markov chain sampling the posterior probability of
the observed 217-3000 GHz thermal dust SED as a function of the two parameters 
$\tilde{I}_{545}$ and $T_2$. More specifically, for each pixel, we are sampling
the posterior given by:

\begin{equation}
\label{eqn:post}
P(\tilde{I}_{545}, T_2|\mathbf{I}) \propto \mathcal{L}(\mathbf{I}|\tilde{I}_{545}, T_2)P(T_2)P(\tilde{I}_{545})
\end{equation}

% be careful about notational consistency with e.g. tildes, etc.. in this sect

Here $\mathbf{I}$ denotes the vector of observed thermal dust intensities 
quoted under the `IRAS convention': $\mathbf{I}$ = ($I_{217}$, $I_{353}$,
$I_{545}$, $I_{857}$, $I_{3000}$). The likelihood function is given by:

%\begin{equation} \label{equ:like}
%\mathcal{L}(\vec{I}~|\tilde{I}_{545}, T_2) = \displaystyle\prod\limits_{i}\mathcal{N}(\tilde{I}_{\nu_{i}}|I_{\nu_{i}}, \sigma_{\nu_i})
%\end{equation}

\begin{equation} \label{equ:like}
\mathcal{L}(\mathbf{I}|\tilde{I}_{545}, T_2) = \textrm{exp}\Big[-\frac{1}{2}(\mathbf{I}-\mathbf{\tilde{I}})\Sigma^{-1}(\mathbf{I}-\mathbf{\tilde{I}})^T\Big]
\end{equation}

%be more careful about notation re: color corrections

Here $\mathbf{\tilde{I}}$ is the vector of predicted observations 
based on Equation \ref{eqn:inten} and the proposed values of $\tilde{I}_{545}$ 
and $T_2$: $\mathbf{\tilde{I}} = $($\tilde{I}_{217}$, 
$\tilde{I}_{353}$, $\tilde{I}_{545}$, $\tilde{I}_{857}$, $\tilde{I}_{3000}$). 
$\Sigma$ is the covariance matrix constructed based on the uncertainties in the
observed intensities:

\begin{equation} \label{equ:covariance}
\Sigma = \begin{pmatrix} \sigma^2_{217}                           & \ldots  & \rho_{217,3000}\sigma_{217}\sigma_{3000} \\ 
                         \vdots                                   & \ddots  & \vdots                                   \\
                         \rho_{3000,217}\sigma_{3000}\sigma_{217} & \ldots  & \sigma^2_{3000}
\end{pmatrix}
\end{equation}

For each pixel $p$ in band $\nu_i$, the variance of the measured value 
$I_{\nu_i}(p)$ is taken to be:

\begin{multline}
\sigma^2_{\nu_i}(p) = c^2_{\nu_i}I_{\nu_i}^2(p) + c^2_{\nu_i}\sigma_{CMB, \nu_i}^2 + (\delta O_{\nu_i})^2 \\
+ n_{\nu_i}^2 + \sigma^2_{CO, \nu_i}(p)  + \sigma^2_{CIBA, \nu_i}
\end{multline}

% is I_nu on RHS inclusive of offsets or is it the ``pure'' thermal dust
% intensity

% need to make sure to cite planckcalib paper

This error budget is modeled after \cite{planckdust} Equation 7, but with some 
modifications and additions. The first term accounts for the multiplicative 
uncertainty on the input maps. Table \ref{table:offs} lists the multiplicative 
calibration uncertainty $c_{\nu}$ for each band. These values are taken from 
Table 11 of \cite{planckcalib}. The second term represents an uncertainty due 
to our subtraction of the \verb|SMICA| CMB model. The analogous term in 
\cite{planckdust} Equation 7 is ($c_{\nu}$$\times$\verb|SMICA|($p$))$^2$, i.e. 
an uncertainty proportional to the CMB model amplitude in each pixel. Because 
this term's spatial dependence can imprint the CMB anisotropies on the derived 
parameters, we have chosen to replace \verb|SMICA|($p$) with a spatially 
constant, RMS value for the CMB amplitude, $\sigma_{CMB, \nu_i}$. 
$\delta O_{\nu_i}$ represents the uncertainty in the band $\nu_i$ zero level 
offset, and the values of $\delta O_{\nu_i}$ can be read off from the second 
column of Table \ref{table:offs}. $n_{\nu_i}$ represents the instrumental noise
in band $\nu_i$. Because using per-pixel noise estimates based on the \PLANCK~
\verb|ii_cov| parameter can imprint features of the survey pattern onto the 
derived parameters, we have adopted a conservative, spatially constant value of
$n_{\nu_i}$ for each band. These values of $n_{\nu_i}$ are listed in Table 
\ref{table:offs}. The next term accounts for the uncertainty on the CO emission
correction, taking $\sigma_{CO, \nu_i}$=0.15$\times$$I_{CO, \nu_i}$ (see 
$\S$\ref{sec:mole}, specifically Equation \ref{equ:molcorr}).

Finally, we include a term to account for the RMS amplitude of the CIBA in
band $\nu_i$, $\sigma_{CIBA, \nu_i}$. The values for the CIBA RMS amplitudes 
are obtained by assuming a $T$=18.3 K, $\beta$=1.0 MBB spectrum for the CIB, 
with 857 GHz normalization from \cite{ciba}. The CIBA not only contributes to 
the per-band variance $\sigma^2_{\nu_i}$, but also to the inter-frequency 
covariances; this is why we have included the off-diagonal terms in the
covariance matrix of Equation \ref{equ:covariance}. In our noise model, the 
CIBA is the only source of inter-frequency covariance. Thus, the off-diagonal 
covariance matrix element between bands $\nu_i$ and $\nu_j$ is given by:

\begin{equation}
\Sigma_{ij} = \rho_{\nu_i, \nu_j}\sigma_{\nu_i}\sigma_{\nu_j} = \rho_{CIBA, \nu_i, \nu_j}\sigma_{CIBA, \nu_i}\sigma_{CIBA, \nu_j}
\end{equation}

With values for $\rho_{CIBA, \nu_i, \nu_j}$ from \cite{covarciba}. The approach
we have taken in accounting for the CIBA is similar to that of 
\cite{planckdust}, Appendix B, in that we treat the CIBA amplitude in each 
pixel as a Gaussian random draw. However, instead of performing a separate 
analysis to gauge the uncertainty on derived dust parameters due to the CIBA, 
we allow the CIBA covariance to propagate naturally into our uncertainties via 
the likelihood function. Still, our treatment of the CIBA is a major 
oversimplification; a more sophisticated approach that accounts for the 
detailed CIBA spatial structure, or even removes the CIBA by subtraction would
be preferable.

% mention that each of the error terms is always rescaled properly according
% to smoothing/rebinning and relevant power spectrum

We include the following prior on the hot dust temperature:

\begin{equation} \label{equ:t2prior}
P(T_2) = \mathcal{N}(T_2|\bar{T}_2, \sigma_{\bar{T}_2})
\end{equation}

With $\bar{T}_2$ = 15.7 K and $\sigma_{\bar{T}_2}$ = 1.4 K. The $T_2$ prior 
mean is chosen based on the typical high-latitude $T_2$ value derived from
the correlation slope analysis of $\S$\ref{sec:global}. We find, as 
desired, that this relatively broad $T_2$ prior has little influence on the 
derived temperatures, other than to regularize the rare pixels with one or more
defective intensities which might otherwise yield unreasonable parameter
estimates. In principle, there can also be an informative prior on 
$\tilde{I}_{545}$. However, we have chosen to assume a uniform prior on the SED
normalization and, as a matter of notation, will omit $P(\tilde{I}_{545})$ 
henceforward. In practice we always perform computations using logarithms of 
the relevant probabilities.

\begin{figure}
\begin{center}
\epsfig{file=tcomparison.eps, width=3.3in}
\caption{\label{fig:comparison} Comparison of temperature maps based on FIR 
dust emission over a $10.5^{\circ}\times8.3^{\circ}$ region centered about 
$(l,b) = (111.6^{\circ}, 18.3^{\circ})$. Top: SFD temperature map based on 
DIRBE $100\mu$m and $240\mu$m, with $\sim$1.3$^{\circ}$ resolution. Center: 
$6.1'$ resolution two-component temperature based on \PLANCK~217-857 GHz and 
SFD 100$\mu$m. Bottom: \cite{planckdust} single-MBB temperature map based on 
\PLANCK~353-857 GHz and 100$\mu$m data, with $5.1'$ FWHM. Both temperature maps
incorporating \PLANCK~observations clearly show a major improvement in angular 
resolution relative to SFD.}
\end{center}
\end{figure}

For each pixel, we initialize the Markov chain with parameters 
$\tilde{I}_{545}$ = $I_{545}$ and $T_2$ consistent with the FDS99 
DIRBE 100$\mu$m/240$\mu$m ratio map $\mathscr{R}$. The initial proposal 
distribution is a two-dimensional normal distribution, with 
$\sigma_{T_2}$=0.25 K, 
$\sigma_{\tilde{I}_{545}}$=max(0.01$\times$$I_{545}$, 0.05 MJy/sr) and
$\rho_{T_2, \tilde{I}_{545}}$=0. We run 5 iterations of burn-in, each 
consisting of 500 MH steps. After each burn-in iteration, we rescale the 
proposal distribution so as to ultimately attain an acceptance fraction 
$f_{acc}$ as close as possible to the optimal value $f_{opt}=0.234$. This is 
accomplished by multiplying the proposal distribution standard deviations by 
$f_{acc}$/$f_{opt}$.

After burn-in, we estimate the parameters and their uncertainties by performing
10,000 sampling steps, with $T_{2, j}$ and $\tilde{I}_{545, j}$ denoting the 
proposed parameter values at the $j^{th}$ step since the end of burn-in. From 
these 10,000 samples, we compute estimates of each parameter's mean, 
$\langle T_2 \rangle$ = $\langle T_{2, j} \rangle$, 
$\langle \tilde{I}_{545} \rangle$ = $\langle \tilde{I}_{545, j} \rangle$, of 
each parameter's variance, 
$\sigma^2_{T_2}$ = $\langle T^2_{2, j} \rangle-\langle T_{2, j} \rangle ^2$, 
$\sigma^2_{\tilde{I}_{545}}$=$\langle \tilde{I}^2_{545, j} 
\rangle-\langle \tilde{I}_{545, j} \rangle ^2$ and of the covariance 
$\sigma_{T_2}\sigma_{\tilde{I}_{545}}$=$\langle T_{2,j}-\langle T_2 \rangle 
\rangle \langle \tilde{I}_{545, j} - \langle \tilde{I}_{545} \rangle \rangle$.

After obtaining this initial estimate of the covariance matrix for each pixel, 
we re-run a second iteration of the entire MCMC procedure, starting from the 
first burn-in period. On this iteration, for each pixel, we begin with a 
proposal distribution that is a two-dimensional Gaussian with covariance equal 
to the first-pass covariance estimate. This gives the each pixel's proposal 
distribution approximately the `right shape', whereas on the first pass we 
started by simply guessing the relative widths of the proposal distribution in 
$\tilde{I}_{545}$, $T_2$, and also assumed that the 
$\rho_{T_2, \tilde{I}_{545}}$=0.

We validated the recovered parameters and uncertainties from our MCMC procedure
by comparing with results based on gridded posterior calculations. These 
comparisons verified that the proposal distribution rescaling and reshaping 
steps that we employ do improve the accuracy of the recovered 
parameters/uncertainties, and that the parameters/uncertainties ultimately 
derived are highly reliable.

Lastly, during post burn-in sampling, we also estimate the monochromatic 
two-component intensity at 545 GHz, $M_{545}$ = $\langle M_{545, j} \rangle$ = 
$\langle \tilde{I}_{545, j}/b_{545}(T_{2,j}) \rangle$, its variance, and the 
545 GHz optical depth $\tau_{545}$=$\langle \tau_{545, j} \rangle$ = 
$\langle M_{545, j}/S_{545, j} \rangle$ and its variance. $\tau_{545}$ and 
$M_{545}$ are more readily useful than the sampling parameters themselves for 
translating our fit results into predictions of reddening ($\S$\ref{sec:ebv}) 
and thermal dust emission ($\S$\ref{sec:lofreq}), respectively. At high
Galactic latitude, we find a typical $T_2$ uncertainty of 0.45 K, and
typical $\tilde{I}_{545}$ fractional uncertainty of 13\%. Figure \ref{fig:sed}
shows a summary of the observed SED and resulting best-fit two-component
model for a single $N_{side}$=2048 pixel, and Figure \ref{fig:post} 
shows posterior PDFs for three pixels. Figure \ref{fig:comparison} shows
a map of our derived hot dust temperature at full-resolution, for a patch
of sky in the Polaris flare region.

\subsection{Low-resolution Fits}
\label{sec:lores}

\begin{figure}
\begin{center}
\epsfig{file=f1_1deg.eps, width=3.3in}
\caption{\label{fig:f1} $1^{\circ}$ FWHM full-sky map of $f_1$ derived from our
low-resolution fits described in $\S$\ref{sec:lores}. Red coloring masks pixels
with appreciable molecular emission, as defined in $\S$\ref{sec:mole}. Such 
pixels should not be trusted in this analysis, which is sensitive to the SED 
shape at low frequencies affected by CO line emission. Variations in $f_1$
along the ecliptic plane are spurious results of imperfect zodiacal light
subtractions. However, interesting astrophysical variations of $f_1$ 
are evident, particularly the trend of increasing $f_1$ with decreasing
Galactic latitude, the relatively low $f_1$ values in the Polaris flare region,
and the clouds with relatively high $f_1$ near the north Galactic pole.}
\end{center}
\end{figure}

As mentioned in $\S$\ref{sec:modeling}, the combination of high S/N and high 
angular resolution afforded by the \PLANCK~HFI maps provides us with the 
opportunity to allow additional parameters of the two-component model, 
previously fixed by FDS99, to vary spatially. Specifically, we consider
allowing $f_1$ to vary, while maintaining $\beta_1$, $\beta_2$, and $q_1/q_2$
spatially constant. In principle, we could alternatively introduce a third free
parameter by permitting $\beta_1$, $\beta_2$ or $q_1/q_2$ to vary while holding
$f_1$ fixed. However, a model in which $f_1$ varies continuously from one line 
of sight to another is the most natural three-parameter scenario, in that $f_1$
variation can be attributed to continuous changes in the dust species' mass
fractions, whereas continuous variations in the other global parameters, which 
represent grain emission/absoprtion properties, seem less plausible.

In order for our variable $f_1$ fits to remain sufficiently constrained 
following the introduction of a third free parameter, we enhance per-pixel S/N 
by smoothing the input maps to $1^{\circ}$ FWHM, and pixelize at $N_{side}=64$.
To best constrain the model parameters in each pixel, we also include 
\PLANCK~100 GHz and 143 GHz, and DIRBE 140$\mu$m and 240$\mu$m, all at 
$1^{\circ}$ resolution.

We now run Markov chains sampling in all three of $f_1$, $\tilde{I}_{545}$ and 
$T_2$, with the posterior given by:

\begin{equation}
\label{eqn:f1post}
P(\tilde{I}_{545}, T_2, f_1|\mathbf{I}) \propto \mathcal{L}(\mathbf{I}|\tilde{I}_{545}, T_2, f_1)P(T_2)P(f_1)
\end{equation}

The likelihood here is conceptually the same as that of Equation 
\ref{equ:like}, but now depends on $f_1$, which can vary from proposal to 
proposal within each individual pixel. The other difference is that 
$\mathbf{I}$ and $\mathbf{\tilde{I}}$ now include 100 GHz, 
143 GHz, 140$\mu$m and 240$\mu$m, in addition to the five bands used for the 
full-resolution fits.

The prior $P(T_2)$ from Equation \ref{equ:t2prior} 
remains unchanged. We adopt the following prior on $f_1$:

\begin{equation} \label{equ:f1prior}
P(f_1) = \mathcal{N}(f_1|\bar{f}_1, \sigma_{\bar{f}_1})
\end{equation}

With $\bar{f}_1$=0.0485 (from Table \ref{tab:global}, model 2) and 
$\sigma_{\bar{f}_1}$=0.005. This is a fairly stringent prior, but we must 
restrict the fit from wandering with too much freedom, as we are attempting to 
constrain three parameters using an SED with only nine intensity measurements, 
several of which are quite noisy.

%%%\begin{figure}
%%%\begin{center}
%%%\epsfig{file=f1_lambert.eps, width=3.3in}
%%%\caption{\label{fig:f1lambert} The results of our low-resolution fit with $f_1$
%%%allowed to vary, shown in Lambert projection to highlight the salient features 
%%%near the north Galactic pole.}
%%%\end{center}
%%%\end{figure}

The resulting full-sky map of $f_1$ is shown in Figure \ref{fig:f1}. A 
general trend of increasing $f_1$ towards lower Galactic latitudes is 
apparent. The other most salient features are the relatively low
values of $f_1$ in the Polaris flare, and the relatively high $f_1$ values
near the north Galactic pole.

%Some of 
%the most striking features occur near the north Galactic pole, and to 
%highlight these regions of unusually high $f_1$, we display the northern
%Galactic hemisphere in Lambert projection in Figure \ref{fig:f1}.

% add comment about high f_1 regions near NGP
% also add sentences referring reader to the two f_1 figures

\subsection{Global Parameters Revisited}
\label{sec:hier}

The posterior sampling framework thus far described also affords us an
opportunity to evaluate the goodness-of-fit for competing dust SED models, and 
thereby cross-check the conclusions of our correlation slope analysis in
$\S$\ref{sec:global}. The basic idea will be to continue evaluating the 
posterior of Equation \ref{eqn:post}, but at low resolution ($N_{side}$=64), 
including all HFI bands as well as DIRBE 100$\mu$m, 140$\mu$m and 240$\mu$m, 
and switching to a uniform prior on $T_2$. Under these circumstances, the 
chi-squared corresponding to the best-fit parameters for each pixel, termed 
$\chi^2_p$, is simply $-2 \times log(P_{max})$. We will refer to the
per-pixel chi-squared per degree of freedom as $\chi^2_{p, \nu}$.

\begin{figure}
\begin{center}
\epsfig{file=chi2_dirbe.eps, width=3.3in}
\caption{\label{fig:chi2_dirbe} Comparison of goodness-of-fit, 
$\chi^2_{\nu}$=$\langle \chi^2_{p, \nu} \rangle$, for various dust SED models,
as described in $\S$\ref{sec:hier}. For single-MBB models with spatially 
constant $\beta$, we varied $\beta$ between 1 and 2 (horizontal axis), 
achieving reduced chi-squared $\chi^2_{\nu}$ shown by the black line, with 
$\beta$=1.57 providing the best fit. Horizontal lines indicate $\chi^2_{\nu}$ 
for other dust emission models considered, including the FDS99 best-fit 
two-component model (Table \ref{tab:global}, model 1, red) and the 
\cite{planckdust} single-MBB model (green). The minimum $\chi^2_{\nu}$ is 
achieved with two-componenent `model 2' from Table \ref{tab:global} (magenta).}
\end{center}
\end{figure}

Because we seek to compare the goodness-of-fit for various dust SED models in 
the diffuse ISM, we restrict to a set of $\sim$10,800 pixels ($\sim$22\% of the
sky), with $|b|>30^{\circ}$ and $|\beta|>10^{\circ}$. We also avoid the SMICA 
inpainting mask, pixels with appreciable CO contamination, and compact sources.
The goodness-of-fit `objective function' we employ to judge the quality of 
a particular dust SED model is $\langle \chi^2_{p, \nu} \rangle$, where the 
average is taken over the aforementioned set of $\sim$10,800 pixels. 
$\langle \chi_{p, \nu}^2 \rangle$ is also equivalent to the reduced 
chi-squared, $\chi^2_{\nu}$, when considering the total number of free 
parameters to be the number of pixels multiplied by the number of free 
parameters per pixel (and similarly for the total number of constraints), and
taking $\chi^2$=$\sum\chi^2_{p}$.

We calculate $\chi^2_{\nu}$ for various dust SED models, independently 
minimizing each $\chi^2_p$ by finding pixel $p$'s best-fitting dust 
temperature and normalization, then evaluating $\langle \chi_{p}^2 \rangle$. 
First, we consider single-MBB models with $\beta$ spatially constant (see the
black line in Figure \ref{fig:chi2_dirbe}). $\beta$=1.57 yields the best fit, 
with $\chi^2_{\nu}$=2.51. This result is in excellent agreement with that of 
$\S$\ref{sec:global}, where we found the best-fit single-MBB model to have 
$\beta$=1.59.

We also evaluated $\chi^2_{\nu}$ for single-MBB models in which
$\beta$ varies spatially. In these cases, we adopted the 0.5$^{\circ}$ 
resolution $\beta$ map from \cite{planckdust}. We started by calculating 
$\chi^2_{\nu}$ using the \cite{planckdust} temperature map, finding 
$\chi^2_{\nu}$=4.68. Note that in this case no per-pixel chi-squared 
minimization was involved, as we simply evaluated $\chi^2_{p}$ for each pixel 
based on the fully-specified \cite{planckdust} emission model. Next, we tested 
a single-MBB model for which we adopted the \cite{planckdust} $\beta$ map, but 
allowed the per-pixel temperature and normalization to vary so as to minimize 
$\chi^2_p$. In this case, we found $\chi^2_{\nu}$=2.51, effectively identical 
to the value found for the spatially constant $\beta$=1.57 single-MBB model. 
This is perhaps unsurprising, as the average $\beta$ value from 
\cite{planckdust} over the mask in question is $\langle\beta\rangle$=1.58. 
This result does suggest, however, that the half-degree variations in $\beta$ 
are not materially improving the goodness-of-fit over the full frequency range 
100-3000 GHz relative to a model with appropriately chosen spatially constant 
$\beta$.

% mention that when t floats, we find a lower T by ~ 1 degree than the Planck 
%team?

We move on to evaluate two-component models, first calculating $\chi^2_{\nu}$ 
with the FDS99 global parameters (Table \ref{tab:global}, model 1). We find
$\chi^2_{\nu}$=2.33, a slight improvement relative to the best-fitting
single-MBB models. Finally, we calculate $\chi^2_{\nu}$ for Table 
\ref{tab:global} model 2, the two-component model favored by our \PLANCK+DIRBE 
correlation slopes. In this case, we achieve the best goodness-of-fit out of 
all the models we have tested, with $\chi^2_{\nu}$=2.11.

Thus, our degree-resolution goodness-of-fit analysis has generally confirmed
the conclusions of $\S$\ref{sec:global}. We find the single-MBB $\beta$ value 
favored by the combination of \PLANCK~and DIRBE to be nearly identical here 
($\beta$=1.57) versus in $\S$\ref{sec:global} ($\beta$=1.59). As in 
$\S$\ref{sec:global}, we also find that the \PLANCK+FIRAS and \PLANCK+DIRBE 
best-fit two-component models from Table \ref{tab:global} outperform single-MBB
alternatives, though only by a relatively small margin in $\chi^2_{\nu}$. 
Still, because our present analysis has $\sim$75,500 degrees-of-freedom, 
$\Delta \chi^2_{\nu}$=0.4 formally corresponds to an enormously significant
improvement in $\chi^2$. The agreement between our correlation slope analysis 
and the present goodness-of-fit analysis is especially encouraging for two main
reasons: (1) in the present analysis, dust temperature has been allowed to vary
on degree scales, whereas in $\S$\ref{sec:global} we assumed a single global 
dust temperature and (2) in the present analysis, our zero-level offsets factor
into the dust temperature, whereas in $\S$6 this was not the case, meaning the 
former and latter analyses agree in spite of their potential to be affected by 
rather different systematics.

% say that, becuase such little improvement for two-component versus
% single MBB in this analysis, we include a single-MBB beta = 1.57 fit in
% our data release ???

% could compare temperatures from correlation slope analysis versus here, but 
% masks are different, so i guess it's just not worth it

\section{Optical Reddening}
\label{sec:ebv}

While the temperature and optical depth maps thus far derived are
useful for making thermal dust emission foreground predictions, estimating 
optical reddening/extinction is another important application of the 
$\tau_{545}$ map. Translating our two-component optical depth to reddening
is especially valuable because our $T_2$ map has $\sim$13$\times$ better
angular resolution than the SFD temperature correction, and thus there is 
reason to believe our two-component reddening estimates may be superior to
those of SFD. However, as discussed in $\S$\ref{sec:replace}, we do not yet 
advocate for the wholesale replacement of SFD, and more detailed work is still 
necessary to determine/quantify the extent to which \PLANCK-based dust maps
might improve reddening estimates relative to SFD.

% mention that this is especially true because Planck gives such a huge
% increase in temperature resolution relative to SFD (and possibly reference
% temperature comparison figure as well)

\subsection{Reddening Calibration Procedure}

\begin{figure}
\begin{center}
\epsfig{file=calib_ebv.eps, width=3.3in}
\caption{\label{fig:calib} Linear fit of $E(B-V)_{SSPP}$ as a function of
two-component $\tau_{545}$, illustrating our procedure for calibrating optical 
depth to reddening.}
\end{center}
\end{figure}

We calibrate optical depth to reddening empirically rather than derive a 
relationship between $\tau_{545}$ and reddening by introducing additional 
assumptions about the dust grain physics and size distribution. To achieve this
empirical calibration, we must adopt a set of calibrator objects for which true
optical reddening is known. There are various possibilities at our disposal. 
\cite{planckdust} calibrated their radiance and $\tau_{353}$ maps to $E(B-V)$ 
using broadband Sloan Digital Sky Survey \citep[SDSS;][]{sdss} photometry for a
set of $\sim$10$^5$ quasars. The SFD calibration was originally tied to a 
sample of 384 elliptical galaxies, but was later revised by 
\citet[hereafter SF11]{schlafly11} based on $\sim$260,000 stars with both 
spectroscopy and broadband photometry available from the SEGUE Stellar 
Parameter Pipeline \citep[SSPP,][]{sspp}.

% make argument that, in models with f1 and the beta's constant, the choice
% of frequency at which tau is evaluated does not matter

\begin{figure*}
\begin{center}
\epsfig{file=ebv_resid.eps, width=6.6in}
\caption{\label{fig:resid} (top left) Residuals of $E(B-V)_{2comp}$ relative to
$E(B-V)_{SSPP}$ as a function of $E(B-V)_{SFD}$. The grayscale represents the 
conditional probability within each $E(B-V)_{SFD}$ bin. The central black line 
shows the moving median. The upper and lower black lines represent the moving 
75th and 25th percentiles respectively. (bottom left) Residuals of 
$E(B-V)_{2comp}$ relative to $E(B-V)_{SSPP}$ as a function of hot dust 
temperature $T_2$. (top right) Same as top left, but illustrating the residuals
of $E(B-V)_{mbb}$, our calibration of the \cite{planckdust} $\tau_{353}$ to 
$E(B-V)_{SSPP}$. (bottom right)  Same as bottom left, but showing the 
$E(B-V)_{mbb}$ residuals as a function of the single-MBB dust temperature from 
\cite{planckdust}. The temperature axes always range from the 0.4$^{th}$ 
percentile temperature value to the 99.6$^{th}$ percentile temperature value.}
\end{center}
\end{figure*}

To calibrate our two-component optical depth to reddening, we make use of the 
stellar sample from SF11. Given a library of model stellar atmospheres, the 
spectral lines of these stars can be used to predict their intrinsic optical 
broadband colors. The `true' reddening is then simply the difference between 
the observed $g-r$ color and the $g-r$ color predicted from the spectral lines.
Applying a color transformation then yields `true' $E(B-V)$ values for 
$\sim$260,000 lines of sight. Throughout our SSPP calibration analysis, we 
restrict to the $\sim$230,000 lines of sight with $|b|$$>$20$^{\circ}$ in order
to avoid stars which may not lie behind the full dust column. The calibration 
of two-component optical depth to $E(B-V)$ is performed as a linear regression 
of $E(B-V)_{SSPP}$ versus $\tau_{545}$. $\tau_{545}$ is considered to be the 
independent variable in this regression, as we ultimately wish to predict 
$E(B-V)$ as a function of optical depth, and $\tau_{545}$ has much higher S/N 
than the SSPP $E(B-V)$ estimates.

% MENTION THAT THE TAU TEMPLATE IS INDEPENDENT OF FREQUENCY FOR 
% TWO-COMPONENT MODELS WITH SPATIALLY CONSTANT BETAS, Q1/Q2 AND F1

This regression is illustrated in Figure \ref{fig:calib}. As expected, there 
is a strong linear correlation between $E(B-V)_{SSPP}$ and $\tau_{545}$. The 
conversion factor from $\tau_{545}$ to $E(B-V)$ is 2.46$\times$10$^{3}$. 
Reassuringly, the best-fit offset is quite close to zero, $\sim$0.6 mmag.

Figure \ref{fig:resid} shows the residuals of our $\tau_{545}$-based reddening 
predictions, $E(B-V)_{2comp}$, relative to the corresponding SF11 reddening 
measurements, $E(B-V)_{SSPP}$, as a function of SFD reddening, $E(B-V)_{SFD}$, 
(top left panel) and as a function of hot dust temperature (bottom left panel).
For comparison, the right panels show analogous residual plots, but with 
respect to reddening predictions based on our calibration of the 
\cite{planckdust} 353 GHz optical depth to $E(B-V)_{SSPP}$ , using the same 
regression procedure employed to calibrate $E(B-V)_{2comp}$. We refer to these 
reddening predictions based on the \cite{planckdust} single-MBB model and 
calibrated to the SF11 measurements as $E(B-V)_{mbb}$.

All four residual plots in Figure 12 show systematic problems at some level. 
The most striking systematic trend is the `bending' behavior of the reddening
residuals versus $E(B-V)_{SFD}$ (top panels), with the median residual 
bottoming out near $-10$ mmag at $E(B-V)_{SFD}$$\approx$0.15 mag. This behavior
is common to both $E(B-V)_{2comp}$ and $E(B-V)_{mbb}$, and in fact was first 
noted in the residuals of $E(B-V)_{SFD}$ itself relative to $E(B-V)_{SSPP}$ by 
SF11 (see their Figure 6). Such a bending behavior is troubling because it 
could indicate a nonlinearity common to many FIR reddening predictions based on
column densities inferred from dust emission. Alternatively, because the SF11 
stars are distributed over the sky in a highly non-uniform manner, the bend 
could arise from aliasing of discrepancies particular to certain sky regions 
(e.g. inner vs. outer Galaxy) on to the $E(B-V)_{SFD}$ axis.

% could add a sentence praising the two-component reddening prediction for
% only having issues at the ~<10 mmag level for EBV_SFD ~< 0.4 mag, since
% it could be a lot worse ...

\begin{figure*}
\begin{center}
\epsfig{file=ebv_resid_ecl.eps, width=6.6in}
\caption{\label{fig:resid_ecl} Same as Figure \ref{fig:resid}, but restricting 
to high ecliptic latitude, $|\beta|>20^{\circ}$. In both the top left and top 
right plots, the bending of the reddening residuals as a function of 
$E(B-V)_{SFD}$ seen in Figure \ref{fig:resid} has been eliminated. Further, the
two-component reddening residual temperature dependence (bottom left) has been 
significantly reduced relative to the corresponding trend shown in Figure 
\ref{fig:resid}. For $E(B-V)_{SFD}$$\gtrsim$0.3 mag, the top row plots appear 
noisy because there are an insufficient number of remaining SSPP points of 
comparison.}
\end{center}
\end{figure*}

The obvious culprit for any potential nonlinearity in FIR-based reddening
estimates is a faulty temperature correction. For this reason, we have included
the bottom panels of Figure \ref{fig:resid}, to check for the presence of
a temperature dependence of the reddening residuals. Indeed, in both the
two-component and single-MBB cases there exists some systematic dependence of
the reddening residuals on temperature. For $T_{mbb}$$\gtrsim$19 K, the 
median residual is reasonably flat, but at lower temperatures (the lowest 
temperature $\sim$20\% of the sky), the median shows trends at the $\sim$10 
mmag level. On the other hand, the median residual in the two-component case 
trends downward with increasing $T_2$ over the entire $T_2$ range shown, with a
peak-to-peak amplitude of $\sim$20 mmag.

\subsection{Rectifying the Reddening Residuals}

In this section we describe our attempts to eliminate the systematic
problems in the two-component reddening residuals shown in the left column
of Figure \ref{fig:resid}. We employed two main strategies: (1) recomputing
the two-component $\tau_{545}$ by re-running our Markov chains after modifying 
the input maps and/or changing the particular two-component model adopted and
(2) making spatial cuts to isolate sky regions in which the 
residuals are especially pristine (or especially problematic).

The following is a list of dust model modifications we tested, but which
proved to have little impact on the reddening residual trends as a function 
of either $E(B-V)_{SFD}$ or $T_2$:

\begin{itemize}
\item Varying each of the global two-component model parameters $\beta_1$, 
$\beta_2$, $q_1/q_2$ and $f_1$ individually while holding the others fixed.
\vspace{-3mm}
\item Allowing $f_1$ to vary spatially as in the fits of $\S$\ref{sec:lores}.
\vspace{-2mm}
\item Changing the mean and/or variance of the $T_2$ prior.
\vspace{-6mm}
\item Varying multiple global parameters at a time e.g. both $f_1$ and 
$q_1/q_2$, restricting to regions of parameter space favored by our 
goodness-of-fit analyses described in $\S$\ref{sec:global} and 
$\S$\ref{sec:hier}.
\end{itemize}

We additionally investigated the following spatial cuts which did not resolve 
the dominant problems noted in the reddening residuals:
\begin{itemize}
\item Separating Celestial north and south.
\vspace{-3mm}
\item Separating Galactic north and south.
\vspace{-3mm}
\item Separating inner and outer Galaxy.
\vspace{-3mm}
\item Combining the above two sets of cuts i.e. separating the Galaxy into
quadrants. 
\vspace{-3mm}
\item Combining these spatial cuts with the dust model changes of the previous
list.
\end{itemize}

However, we found that changing the zero level offsets of the input maps 
had a significant effect on the strength of the anticorrelation between median 
reddening residual and $T_2$. In particular, we experimented with perturbing 
the zero level offset of \PLANCK~857 GHz while correspondingly changing the 
zero levels of the remaining \PLANCK~maps based on the prescription of 
$\S$\ref{sec:relzero}. We also experimented with changing the zero level of SFD
\verb|i100|, independent of the other zero levels. Unfortunately, completely
flattening the reddening residual dependence on $T_2$ required unreasonably 
large zero level modifications. For example, flattening the $T_2$ residual 
required adding $\gtrsim$0.6 MJy/sr to the \verb|i100| map. Such an offset is 
implausible, being an order of magnitude larger than the nominal \verb|i100| 
zero level uncertainty quoted by SFD, and comparable to the entire 3000 GHz CIB
monopole signal. Furthermore, we note that even these large zero level 
modifications had virtually no effect in eliminating the reddening residual 
`bend' versus $E(B-V)_{SFD}$. Thus, changing the zero level offsets showed 
hints of promise in rectifying the reddening residual temperature dependence, 
but cannot completely resolve the systematic trends in reddening residuals.
% last few words here are kind of bad, need better terminology

The only solution we have been able to identify that both removes the `bend' 
vs. $E(B-V)_{SFD}$ and simultaneously reduces the temperature dependence of the
reddening residuals is cutting out the ecliptic plane by restricting to 
$|\beta|>20^{\circ}$. In this case, we completely eliminated the bending 
behavior of the residual versus $E(B-V)_{SFD}$, and significantly reduced the 
$T_2$ dependence to a peak-to-peak amplitude of only $\sim$10 mmag (see Figure 
\ref{fig:resid_ecl}). Figure \ref{fig:resid_ecl} still includes the single-MBB 
plots (right column), to show that the bend versus $E(B-V)_{SFD}$ is eliminated
by the $|\beta|$ cut, even for the single-MBB model. However, the single-MBB 
residuals still differ systematically from zero for $T$$\lesssim$$19$ K. 
Perhaps the improvements in the two-component reddening residuals after 
restricting to high ecliptic latitude should come as no surprise, given that 
the ecliptic plane is the most obvious systematic problem with our temperature 
map (see the full-sky results shown in Figure \ref{fig:results}).

After cutting the ecliptic plane, we found that only small zero level
perturbations were required to fully flatten the temperature residuals,
while still maintaining flat residuals versus $E(B-V)_{SFD}$. The optimal
offsets we found were $\pm$0.08 MJy/sr to \verb|i100| and 857 GHz respectively
(see Figure \ref{fig:resid_offs}). These offsets are well within reason, 
given the nominal zero level uncertainties quoted in Table \ref{table:offs}.

\begin{figure}
\begin{center}
\epsfig{file=resid_offs.eps, width=3.3in}
\caption{\label{fig:resid_offs} Two-component reddening residuals after 
restricting to high ecliptic latitude ($|\beta|>20^{\circ}$) and perturbing the
\texttt{i100} and 857 GHz zero levels by $+$0.08 MJy/sr and $-$0.08 MJy/sr 
respectively. The bending behavior as a function of $E(B-V)_{SFD}$ has been
eliminated, and virtually no temperature dependence remains.}
\end{center}
\end{figure}

% but then we tried ECL cuts, and in combination with modest offsets, that 
% worked

%%\section{Systematics}
%%
%%\subsection{CIBA}
%%Perhaps the dominant limitation of our fits is the presence of cosmic
%%infrared background anisotropies (CIBA) bleeding through into our maps of 
%%Galactic dust temperature and optical depth in high-latitude regions 
%%\citep{ciba}.
%%\subsubsection{Simulations}

%%\subsubsection{Optimizing Relative Weights}

%%\subsection{Zero Point}
%%Simulate the effect of getting zero points wrong to quantiy this
%%systematic a bit.

%%\subsection{$T_{2}$ Prior}
%%Use FIRAS data to justify $T_2$ prior. Could I use the DIRBE 100/240 micron
%%ratio to get a $1.3^{\circ}$ resolution prior on $T_2$?

%%\subsection{Other Limitations}
%%\IRAS~is missing in some parts of sky which will therefore not truly have 
%%6.1' resolution.
%%any other miscellaneous issues?

\section{Low-frequency Emission Predictions}
\label{sec:lofreq}
FDS99 originally performed their FIRAS+DIRBE dust SED analysis for the sake of
accurately forecasting low-frequency CMB foregrounds. Recently, Galactic CMB
foregrounds, especially in the 100-150 GHz frequency range, have become a focal
point of cosmology owing to the \cite{bicep2} $B$-mode polarization results. 
Here we show that our two-component foreground predictions are more accurate 
than those based on the \cite{planckdust} single-MBB model over the 100-217 GHz
frequency range.

\begin{figure*}
\begin{center}
\epsfig{file=compare_lofreq.eps, width=6.8in}
\caption{\label{fig:lofreq} Comparison between low-frequency thermal dust 
emission predictions from our best-fit two-component model (Table 
\ref{tab:global}, model 2) and those of \cite{planckdust}. The top row shows
scatter plots of the \cite{planckdust} predictions versus observed \PLANCK~100 
GHz (left), \PLANCK~143 GHz (center) and \PLANCK~217 GHz (right). The bottom 
row shows scatter plots of the corresponding two-component predictions versus
\PLANCK~observations. In all cases, the blue line indicates the best-fit linear
relationship, while the red line represents a perfect match between predictions
and observations. The lines of best-fit illustrate that the single-MBB model 
systematically underpredicts emission (in the multiplicative sense) by 18.8\%, 
12.6\% and 7.9\% at 100, 143 and 217 GHz respectively. On the other hand, by 
the same metric, the two-component model predictions at 100-217 GHz are always 
accurate to within $\le$2.2\%. The two-component fit results shown are based on
217-3000 GHz observations, meaning that the 100 GHz and 143 GHz predictions are
truly extrapolations, while the 217 GHz agreement is enforced by the fitting 
process itself to some extent.}
\end{center}
\end{figure*}

To assess the accuracy of low-frequency emission predictions, we compare
the \PLANCK~HFI map at each of 100, 143, 217 GHz to the corresponding 
single-MBB and two-component predictions, with all maps smoothed to 
$1^{\circ}$ FWHM and binned down to $N_{side}=64$. We restrict to the 
same set of pixels used for the goodness-of-fit analysis of 
$\S$\ref{sec:hier}, with $|b|>30^{\circ}$ and $|\beta|>30^{\circ}$, also 
avoiding molecular emission, the SMICA inpainting mask, and compact sources. We
then perform a linear fit between the \PLANCK~observed emission and the 
predicted emission at each frequency and for each emission model. For these
fits, we consider the predicted emission to be the independent variable, 
since it has higher S/N than the observations, especially at 100 and 143 GHz.
We also assign pixel weights proportional to the predicted emission, so that
the best-fit lines faithfully capture the linear trend exhibited 
without being biased by the large number of very low S/N pixels with minimal 
emission. Scatter plots between the predicted and observed emission are shown 
in Figure \ref{fig:lofreq}. The best-fit lines are overplotted and their 
equations are given in the top left corner of each subplot.

In both the single-MBB and two-component cases, all of the best fit offsets
are within the uncertainties quoted in Table \ref{table:offs}. On the other 
hand, the top row of Figure \ref{fig:lofreq} shows that the \cite{planckdust} 
single-MBB predictions yield slopes substantially different from unity: 1.079 
at 217 GHz, 1.126 at 143 GHz, and 1.188 at 100 GHz. The fact that the slopes 
are larger than unity indicates that the \cite{planckdust} predictions are 
systematically low. The systematic underprediction evidently become gradually 
more severe as lower frequencies are considered, with a 7.9\% underprediction 
at 217 GHz, a 12.6\% underprediction at 143 GHz and an 18.8\% underprediction 
at 100 GHz.

For the case of the two-component model, we perform full-resolution 217-3000 
GHz fits using the \PLANCK+DIRBE favored global parameters (Table 
\ref{tab:global}, model 2), then smooth to $1^{\circ}$ FWHM and bin down to 
$N_{side}=64$ before predicting the 100-217 GHz emission. The bottom row of
Figure 14 shows that each of the best-fit lines is very similar to the 
corresponding red line which represents a perfect match between predicted and 
observed emission. More quantitatively, the two-component slopes are all within
2.2\% of unity: 0.978 at 217 GHz, 0.986 at 143 GHz and 1.022 at 100 GHz. 
We note that at 217 GHz, the good agreement is in some sense predetermined by 
the fact that \PLANCK~217 GHz has been included in our two-component MCMC fits.
On the other hand, the 143 and 100 GHz predictions are based on extrapolation.

We conclude from our predicted versus observed emission comparison that our 
two-component model outperforms the \cite{planckdust} single-MBB model at 
predicting Galactic thermal dust emission in diffuse regions from 100-217 GHz.

% could add a last sentence trying to soften this, acknowledging that 
% planck team warned about their model below 353 GHz

% mention imprint of Planck team's half-degree on low-freq extrapolation ?

% is this the section where, potentially, plots of the residuals might go?

%%%It is true that we have smoothed the input intensity maps for this study 
%%%to $6.1'$ FWHM, which is nominally lower resolution than the $5'$ 
%%%\cite{planckdust} maps. However, as is apparent from figure 
%%%\ref{fig:comparison}, the fact that \cite{planckdust} modeled $\beta$ at 
%%%$0.5^{\circ}$ resolution has effectively blurred their dust temperature map.

\section{Data Release}
\label{sec:release}
We are releasing a set of $N_{side}$=2048 HEALPix maps in Galactic coordinates 
which summarize the results of our full-resolution two-component dust fits. 
Low-resolution renderings of our full-sky dust temperature and optical depth 
maps are shown in Figure \ref{fig:results}. Our data release also includes
software utilities for obtaining emission and reddening predictions from our 
\PLANCK-based two-component fits. Refer to the data release documentation and 
FITS file headers for further details.\footnote{\texttt{http://faun.rc.fas.harvard.edu/ameisner/planckdust}}

% mention that DR includes software to obtain emission and extinction 
% predictions

% if we're not going to be claiming to be superseding SFD in terms of 
% reddening, then do we even want to include anything extinction-related
% (data files and/or code) in this data release ?

%software utilities
%provide maps at multiple different nsides ?
%   -> if so, actually smooth the maps before rebinning, or leave 
%      undersampled?
% provide uncertainty maps for tau and T2 ??

\begin{figure*} [ht]
\begin{center}
\epsfig{file=results.eps, width=7.0in}
\caption{\label{fig:results} (top) Hot dust temperature derived from our 
full-resolution two-component model fits of \PLANCK~217-857 GHz and SFD 
100$\mu$m, downbinned to 27.5$'$ resolution. (bottom) Corresponding full-sky 
map of best-fit 545 GHz optical depth.}
\end{center}
\end{figure*}

\section{Conclusions}
\label{sec:conclusion}

\subsection{Single-MBB versus Two-component emission}
A major aim of this work has been to determine whether the  FDS99 
two-component dust emission model remains favored over single-MBB models when 
swapping the \PLANCK~HFI maps for FIRAS at frequencies below 1250 GHz. We 
compared dust SED models in two ways (1) by fitting a 100-3000 GHz spectrum 
composed of per-band correlation slopes versus \PLANCK~857 GHz (2) by finding 
the best-fit dust temperature and optical depth per line-of-sight, with each 
pixel's SED comprised of 100-3000 GHz \PLANCK+DIRBE data, and comparing the 
average goodness-of-fit under various emission models.

In both the correlation slope analysis of $\S$\ref{sec:global} and the 
goodness-of-fit analysis of $\S$\ref{sec:hier} we found that the 
best-fit \PLANCK+DIRBE two-component model (Table \ref{tab:global}, model 2) 
outperformed the best-fit single-MBB model, but by a lesser margin in 
$\chi^2_{\nu}$ than found by FDS99 using FIRAS+DIRBE. Specifically, our 
best-fit \PLANCK+DIRBE two-component model yielded an improvement of 
$\Delta \chi^2_{\nu}$=3.41 ($\S$\ref{sec:global}) and $\Delta \chi^2_{\nu}$=0.4
($\S$\ref{sec:hier}). This represents a far less dramatic contrast in 
$\chi^2_{\nu}$ than found by the FDS99 correlation slope analysis, 
$\Delta \chi^2_{\nu}$=29.2. Perhaps a relative lack of discrimination amongst 
competing dust SED models when relying on \PLANCK+DIRBE is to be expected, 
given that our constraints include only nine broad frequency channels, whereas 
FDS99 employed $>$200 narrow bands. Still, $\Delta \chi^2_{\nu}$=0.4 from 
$\S$\ref{sec:hier} is formally of enormous significance, given the 
$\sim$75,000 degrees of freedom in that analysis.

% in above paragraph, say which single-MBB beta(s) we find to be favored

Nevertheless, we have established that the two-component emission model
remains viable in light of the \PLANCK~HFI data, and that the FIR/submm 
dust SED's preference for two MBB components rather than just one is not
simply an idiosyncracy of the FIRAS spectra. Furthermore, we showed in 
$\S$\ref{sec:lofreq} that our 100-217 GHz two-component emission predictions
are on average accurate to within 2.2\%, whereas the \cite{planckdust} 
single-MBB model systematically understimates low-frequency dust emission by 
18.8\% at 100 GHz, 12.6\% at 143 GHz and 7.9\% at 217 GHz. We therefore 
recommend that those interested in thermal dust foregrounds in the 100-3000 GHz
frequency range use our data release to predict unpolarized dust emission, at 
the very least in order to help determine the level at which the choice of dust
emission model may influence their conclusions.
%last sentence needs work

% sentence criticizing the half-degree beta single MBB model of Planck team
% in terms of the imprint that leaves on low-frequency emission 
% predictions ?

\subsection{Towards a Replacement for SFD}
\label{sec:replace}
Because of the broad frequency coverage and high angular resolution afforded
by the \PLANCK~HFI full-sky maps, we initially speculated that a \PLANCK~based 
extinction map might easily outperform SFD, the current industry standard data 
product. However, at this point in time, we do not yet recommend that the 
results presented in this work be considered a replacement for SFD in terms of 
optical extinction/reddening estimates.

% CIBA
The CIBA remains a major imperfection that still requires further 
investigation. The CIB anisotropies are very evident in low-dust regions of our
maps of optical depth and predicted dust emission. As described in 
$\S$\ref{sec:mcmc}, we have propagated the CIBA RMS amplitudes and 
inter-frequency covariances into our uncertainty estimates through the 
likelihood function in our MCMC procedure. However, this treatment falls far 
short of actually removing the spatial imprint of the CIBA on our derived 
parameters. The CIB anisotropies are more prominent in our optical depth map 
relative to that of SFD because of the lower-frequency \PLANCK~maps we rely 
upon to achieve a high-resolution temperature correction.

Imperfect zodi corrections represent a second major limitation of our results.
The ecliptic plane's prominence in our full-sky temperature map (Figure 
\ref{fig:results}) suggests that the zodiacal light subtractions performed on 
the input maps are not ideal. Our comparisons of the FIR maps used in this 
study against H\,\textsc{i} emission bear out this notion, further revealing 
that the imperfect zodi corrections are not limited to \verb|i100| (the 
map for which one would expect zodiacal contamination to be most 
problematic), but in fact are noticeable in all of the HFI 
\verb|R1.10_nominal_ZodiCorrected| maps as well. We deemed it infeasible to 
reconsider all of the \PLANCK~zodi corrections in addition to the 3000 GHz zodi
correction as a part of this study. Perhaps the forthcoming \PLANCK~data 
release, including observations acquired during additional orbits, will 
effectively eliminate any residual zodi problems in the \PLANCK~bands. Under 
these circumstances it might then be worthwhile to consider revising the DIRBE 
100$\mu$m zodi correction provided by SFD.

% could add a sentence noting that similar zodi corrections are evident
% in all existing, comparable temperature map data products : Planck 2013a,
% SFD98 and FDS99

% audit for word ``zodi'' ... is this an acceptable jargon term ?
% could use IPD as an abbreviation at certain points, not sure if i've 
% defined that acronym already somewhere else in the text?

Irrespective of the notable imperfections in our results, more detailed
comparisons between our reddening estimates here and those of SFD are required 
to determine/quantify which map is superior in particular applications. One 
definitive improvement of our reddening estimates relative to those of SFD is 
our ability to quote reddening uncertainties, which results from the 
probabilistic framework of $\S$\ref{sec:mcmc}. The extinction estimates 
from this work can also be employed as an alternative to those of SFD, to gauge
the impact of dust map choice in a specific end user's application.

% perhaps zodi will just magically be fixed in the forthcoming Planck
%   ---> do i want a figure illustrating the CIBA problem at high latitude ?
%  ---> zodi
%  ---> more testing needed, in any event, to prove Planck based reddening
%       estimates outperform SFD

% should mention that one definitive improvement of this work relative
% to FDS99 is that there are uncertainty/covariances provided for
% the derived parameters !!!!

% should there be a future work section or can the suggestions/plans of
% future work just be a part of ``SFD replacement'' subsection

% mention here the idea of getting even higher resolution by folding
% in WISE ?

This material is based upon work supported by the National Science Foundation 
Graduate Research Fellowship under Grant No. DGE1144152. We acknowledge support
of NASA grant NNX12AE08G for this research. Based on observations obtained with
Planck (http://www.esa.int/Planck), an ESA science mission with instruments and
contributions directly funded by ESA Member States, NASA, and Canada. This 
research made use of the NASA Astrophysics Data System (ADS) and the IDL 
Astronomy User's Library at Goddard. \footnote{Available at 
\texttt{http://idlastro.gsfc.nasa.gov}}

\bibliographystyle{apj}
\bibliography{twocomp.bib}

\end{document}